### 2017-01-26 paper.Rmd
Integration of cost data and first draft paper

---
title: "Cost-effectiveness of a malaria elimination campaign"
author: "Laia Cirera, Joe Brew, Elisa Sicuri"
date: "26 de enero de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
library(readxl)
library(tidyverse)

# Read in Laias cost data
costs <- read_excel("public_data/costs.xlsx")

# Get yearly malaria cases for magude

# Create a column for % severe malaria
# This will be the number of cases of ...

# Read in incidence data
bes <- read_csv("data/cleaned/cases.csv")
```

# Introduction

The average incidence of malaria for our years of observation was `r round(mean(bes$pk, na.rm=TRUE),digits=2)`

# Methodology

WE analyzed data from bla bla bla

# Results

Our results show that the effectiveness of MALTEM's interventions is as follows:

```{r}
plot(x=costs$year , y=costs$`Case fatality rate`, type="l",
     xlab="YEAR", ylab="rate", main="Case fatality rate ACROSS TIME",
     col="blue")
```

And here is another figure.

```{r, fig.height = 8}

ggplot(data = bes,
       aes(x = date,
           y = pk)) +
  geom_line(aes(color = age)) +
  facet_wrap(~district) +
  labs(x = 'Date',
       y = 'Incidence (cases per 1,000)',
       title = 'Malaria incidence over time by district',
       subtitle = 'Cases per 1,000 inhabitants') +
  scale_color_manual(name = 'Age',
                     values = c('darkgreen', 'darkorange'))
```

# Discussion

### 2017-01-26 combine_bes_data.R

@@ -251,7 +251,7 @@ df_adjusted <-
   mutate(pk = p * 1000)
 
 # Write data
-write_csv(df_adjusted, 'data/cleaned/cases.csv')
+# write_csv(df_adjusted, 'data/cleaned/cases.csv')
 
 # See incidence
 library(cism)

### 2017-01-27 get_weather_data.R
weather data

@@ -1,5 +1,6 @@
 library(raster)
 library(R.utils)
+library(tidyverse)
 
 # Define function for creating link to data
 create_url <- function(date = '2010-01-01'){
@@ -54,3 +55,78 @@ for (i in 1:length(dates)){
     }
   })
 }
+
+# Aggregate weather data
+###########################
+
+# Get a map of maputo province
+library(cism)
+maputo <- moz2
+maputo <- maputo[which(maputo@data$NAME_1 == 'Maputo'),]
+
+# Get the centroid of each district in maputo province
+centroids <-
+  data.frame(district = toupper(maputo@data$NAME_2),
+             x = coordinates(maputo)[,1],
+             y = coordinates(maputo)[,2])
+
+
+# Get a place to store aggregated data
+if(!dir.exists('weather_data/aggregated')){
+  dir.create('weather_data/aggregated')
+}
+
+# Read each of the files
+files <- dir('weather_data/')
+files <- files[grepl('.tif', files, fixed = TRUE)]
+
+# Go into weather data
+setwd('weather_data/')
+
+# Read in each file and combine
+results <- list()
+for (i in 1:length(files)){
+  this_file <- files[i]
+  this_date <- as.Date(gsub('.tif', '', this_file, fixed = TRUE))
+  r <- raster(this_file)
+  
+  # Extract the values
+  x <- raster::extract(r, centroids[,2:3])
+  
+  # Create a dataframe output
+  output <- centroids %>%
+    dplyr::select(district)
+  output$date <- this_date
+  output$precipitation <- x
+  results[[i]] <- output
+  message(this_date)
+}
+
+# Add together all the results
+precipitation <- bind_rows(results)
+precipitation <- 
+  precipitation %>%
+  dplyr::select(date, district, precipitation)
+precipitation$year <- as.numeric(format(precipitation$date, '%Y'))
+precipitation$month <- as.numeric(format(precipitation$date, '%m'))
+precipitation$day <- as.numeric(format(precipitation$date, '%d'))
+precipitation$week <- as.numeric(format(precipitation$date, '%U')) + 1
+
+
+# Aggregate
+precipitation_daily <- precipitation
+precipitation_weekly <- 
+  precipitation %>%
+  group_by(year, week, district) %>%
+  summarise(precipitation = sum(precipitation, na.rm = TRUE))
+precipitation_monthly <- 
+  precipitation %>%
+  group_by(year, month, district) %>%
[…]

### 2017-01-27 master.R

@@ -1,0 +1,13 @@
+library(tidyverse)
+df <- read_csv('data/cleaned/cases.csv')
+precip <- read_csv('weather_data/aggregated/precipitation_weekly.csv')
+
+df$month <- as.numeric(format(df$date, '%m'))
+df$day <- as.numeric(format(df$date, '%d'))
+
+df <-
+  left_join(x = df,
+            y = precip,
+            by = c('district', 'year', 'week'))
+
+write_csv(df, 'data/cleaned/cases_with_precipitation.csv')

### 2017-01-27 plot_weather_data.R

@@ -9,5 +9,5 @@ for (i in 1:length(files)){
   plot(r, 
        main = this_date)
   message(this_date)
-  Sys.sleep(2)
+  Sys.sleep(0.5)
 }

### 2017-01-27 get_weather_data.R
done with weather data integration

@@ -29,7 +29,6 @@ if(!dir.exists('weather_data')){
 for (i in 1:length(dates)){
   this_date <- dates[i]
   start_time <- Sys.time()
-  message(this_date)
   try({
     # Define a file name
     file_name <- paste0('weather_data/', this_date, '.tif')
@@ -59,16 +58,37 @@ for (i in 1:length(dates)){
 # Aggregate weather data
 ###########################
 
-# Get a map of maputo province
-library(cism)
-maputo <- moz2
-maputo <- maputo[which(maputo@data$NAME_1 == 'Maputo'),]
+## Get a map of maputo province
+# library(cism)
+# maputo <- moz2
+# maputo <- maputo[which(maputo@data$NAME_1 == 'Maputo'),]
+# centroids <-
+#   data.frame(district = toupper(maputo@data$NAME_2),
+#              x = coordinates(maputo)[,1],
+#              y = coordinates(maputo)[,2])
 
 # Get the centroid of each district in maputo province
-centroids <-
-  data.frame(district = toupper(maputo@data$NAME_2),
-             x = coordinates(maputo)[,1],
-             y = coordinates(maputo)[,2])
+if('coordinates_of_maputo_districts.csv' %in% dir('public_data')){
+  centroids <- read_csv('public_data/coordinates_of_maputo_districts.csv')
+} else {
+  library(ggmap)
+  centroids <- data_frame(district = c('BOANE',
+                                       'MAGUDE',
+                                       'MANHICA',
+                                       'MARRACUENE',
+                                       'MATOLA',
+                                       'MATUTUINE',
+                                       'MOAMBA',
+                                       'NAMAACHA'))
+  centroids$y <- centroids$x <- NA
+  for (i in 1:nrow(centroids)){
+    coords <- geocode(location = paste0(centroids$district[i], 
+                                        ' DISTRICT, MOZAMBIQUE'))
+    centroids$x[i] <- coords[,1]
+    centroids$y[i] <- coords[,2]
+    write_csv(centroids, 'public_data/coordinates_of_maputo_districts.csv')
+  }
+}
 
 
 # Get a place to store aggregated data
@@ -82,51 +102,66 @@ files <- files[grepl('.tif', files, fixed = TRUE)]
 
 # Go into weather data
 setwd('weather_data/')
-
-# Read in each file and combine
-results <- list()
-for (i in 1:length(files)){
[…]

### 2017-01-27 master.R

@@ -11,3 +11,33 @@ df <-
             by = c('district', 'year', 'week'))
 
 write_csv(df, 'data/cleaned/cases_with_precipitation.csv')
+
+library(cism)
+ggplot(data = df,
+       aes(x = date,
+           y = precipitation)) +
+  geom_line(alpha = 0.8,
+            color = 'darkgreen') +
+  facet_wrap(~district) +
+  labs(x = 'Date',
+       y = 'Milimeters',
+       title = 'Rainfall',
+       subtitle = 'Province of Maputo') +
+  theme_cism()
+
+
+cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(df$year)))
+ggplot(data = df %>%
+         mutate(day_of_year = as.numeric(format(date, '%j'))),
+       aes(x = day_of_year,
+           y = precipitation)) +
+  geom_line(alpha = 0.8,
+            aes(color = factor(year))) +
+  facet_wrap(~district) +
+  labs(x = 'Day of year',
+       y = 'Milimeters',
+       title = 'Rainfall',
+       subtitle = 'Province of Maputo') +
+  theme_bw() +
+  scale_color_manual(name = 'Year',
+                     values = cols)

### 2017-01-31 get_bes_data.R
reorganization of files, integration of weather data

@@ -85,8 +85,8 @@ df <- df %>%
 # Get a "date" column (approximate)
 df$date <- as.Date(paste0(df$year, '-01-01')) + (7 * (df$week -1))
 
-df_adjusted <- df
-df_adjusted$cases <- df_adjusted$value
+df <- df
+df$cases <- df$value
 
 
 # # Get the overlap period
@@ -144,14 +144,14 @@ df_adjusted$cases <- df_adjusted$value
 #   mutate(cases = value / d_over_m) %>%
 #   dplyr::select(-d_over_m) %>%
 #   filter(!id %in% unique(top$id))
-# df_adjusted <-
+# df <-
 #   bind_rows(top,
 #             bottom) %>%
 #   dplyr::select(-id, -overlap)
 
 # Reorder columns
-df_adjusted <-
-  df_adjusted %>%
+df <-
+  df %>%
   dplyr::select(date,
                 year,
                 week,
@@ -238,123 +238,133 @@ pop <- pop %>%
 pop$district[pop$district == 'MANHIÇA'] <- 'MANHICA'
 
 # Join population to data
-df_adjusted  <- df_adjusted %>%
+df  <- df %>%
   left_join(pop,
             by = c("year", "district", 'age'))
 
 
 
 # Make incidence
-df_adjusted <- 
-  df_adjusted %>%
+df <- 
+  df %>%
   mutate(p = cases / population) %>%
   mutate(pk = p * 1000)
 
+# Get some date objects
+df$month <- as.numeric(format(df$date, '%m'))
+df$day <- as.numeric(format(df$date, '%d'))
+
+
 # Write data
-# write_csv(df_adjusted, 'data/cleaned/cases.csv')
+write_csv(df, 'data/outputs/cases.csv')
 
-# See incidence
-library(cism)
-ggplot(data = df_adjusted,
-       aes(x = date,
-           y = pk)) +
-  geom_line(aes(color = age)) +
-  facet_wrap(~district) +
-  labs(x = 'Date',
-       y = 'Incidence (cases per 1,000)',
-       title = 'Malaria incidence over time by district',
-       subtitle = 'Cases per 1,000 inhabitants') +
-  scale_color_manual(name = 'Age',
-                     values = c('darkgreen', 'darkorange')) +
-  theme_cism()
+# Remove unecessary objects
+rm(d2, mb, pop,
+   final_list, i, sh,
+   read_population)
 
-# See chart fo reach district
[…]

### 2017-01-31 get_irs_data.R

@@ -1,0 +1,4 @@
+library(tidyverse)
+
+irs <- read_delim('data/IRS_Maputo Province_clean.csv',
+                  delim = ';')

### 2017-01-31 get_weather_data.R

@@ -165,3 +165,8 @@ if(file.exists('aggregated/precipitation_daily.csv') &
 }
 
 setwd('..')
+
+# Remove unecessary objects
+rm(centroids,
+   dates, file_name, files, i,
+   start_time, this_date, create_url)

### 2017-01-31 master.R

@@ -1,16 +1,37 @@
+# Libraries
 library(tidyverse)
-df <- read_csv('data/cleaned/cases.csv')
-precip <- read_csv('weather_data/aggregated/precipitation_weekly.csv')
 
-df$month <- as.numeric(format(df$date, '%m'))
-df$day <- as.numeric(format(df$date, '%d'))
+# BES data
+source('get_bes_data.R')
 
+# Precipitation data
+source('get_weather_data.R')
+
+# IRS data
+source('get_irs_data.R')
+
+# Join malaria cases with precipitation
 df <-
   left_join(x = df,
-            y = precip,
+            y = precipitation_weekly,
             by = c('district', 'year', 'week'))
 
-write_csv(df, 'data/cleaned/cases_with_precipitation.csv')
+# Join with IRS
+df <- 
+  left_join(x = df,
+            y = irs %>%
+              dplyr::select(-houses) %>%
+              rename(people_irs = people),
+            by = c('district', 'year')) 
+
+# Get IRS coverage
+df <- df %>%
+  group_by(district, year, week) %>%
+  mutate(district_population = sum(population, na.rm = TRUE)) %>%
+  mutate(irs_coverage = people_irs / district_population * 100)
+
+
+write_csv(df, 'data/outputs/master.csv')
 
 library(cism)
 ggplot(data = df,
@@ -41,3 +62,16 @@ ggplot(data = df %>%
   theme_bw() +
   scale_color_manual(name = 'Year',
                      values = cols)
+
+
+# IRS coverage
+x <- df %>%
+  group_by(district, year) %>%
+  summarise(irs_coverage = mean(irs_coverage))
+
+ggplot(data = x,
+       aes(x = year, 
+           y = irs_coverage,
+           color = district,
+           group = district)) +
+  geom_line()

### 2017-02-01 get_irs_data.R
changes

@@ -1,4 +1,4 @@
 library(tidyverse)
+library(readxl)
 
-irs <- read_delim('data/IRS_Maputo Province_clean.csv',
-                  delim = ';')
+irs <- read_excel('data/IRS_Maputo Province_clean.xlsx')

### 2017-02-01 master.R

@@ -28,10 +28,119 @@ df <-
 df <- df %>%
   group_by(district, year, week) %>%
   mutate(district_population = sum(population, na.rm = TRUE)) %>%
-  mutate(irs_coverage = people_irs / district_population * 100)
+  ungroup %>%
+  mutate(irs_coverage = people_irs / district_population * 100) %>%
+  # Remove unecessary variables
+  dplyr::select(-people_irs, -district_population)
 
-
+# Save a copy
 write_csv(df, 'data/outputs/master.csv')
+
+# Generate a wide dataset
+if(file.exists('data/outputs/master_wide.csv')){
+  df_wide <- read_csv('data/outputs/master_wide.csv')
+} else {
+  df_wide <- df
+  
+  districts <- sort(unique(df$district))
+  years <- sort(unique(df$year))
+  
+  # Generate irs lag
+  df_wide$irs_coverage_last_year <- NA
+  for (i in 1:nrow(df)){
+    this_year <- df$year[i]
+    this_district <- df$district[i]
+    df_wide$irs_coverage_last_year[i] <-
+      df_wide$irs_coverage[df_wide$district == this_district &
+                        df_wide$year == this_year - 1][1]
+  }
+  
+  # Generate precipitation variables
+  starter <- seq(120, 1, by = -10)
+  ender <- seq(1, 120, by = 10)
+  
+  # Create empty columns
+  for (s in starter){
+    for (e in ender){
+      if(s >= e){
+        df_wide[,paste0('precipitation_from_',
+                   s, '_to_',
+                   e, '_days_ago')] <- 
+          NA
+        df_wide[,paste0('max_precipitation_from_',
+                   s, '_to_',
+                   e, '_days_ago')] <- 
+          NA
+        df_wide[,paste0('min_precipitation_from_',
+                   s, '_to_',
+                   e, '_days_ago')] <- 
+          NA
+        df_wide[,paste0('rainy_days_from_',
+                   s, '_to_',
+                   e, '_days_ago')] <- 
+          NA
+        df_wide[,paste0('dry_days_from_',
+                   s, '_to_',
+                   e, '_days_ago')] <- 
+          NA
+      }
+    }
+  }
+  
+  # Populate emtpy columns
+  for (s in starter){
+    for (e in ender){
+      message(paste0(s, '  ', e))
+      if(s >= e){
+        for (i in 1:nrow(df_wide)){
+          this_district <- df_wide$district[i]
[…]

### 2017-02-01 paper.Rmd
proves tontes
Finished and adapted costs inputs to be able to do data analysis and calculations through R
paper main structure and some text

@@
 author: "Laia Cirera, Joe Brew, Elisa Sicuri"
-date: "26 de enero de 2017"
+date: "1 de fevereiro de 2017"
 output: html_document
@@
 
-# Read in Laias cost data
+# Read in costs database 
 costs <- read_excel("public_data/costs.xlsx")
 
-# Get yearly malaria cases for magude
-
-# Create a column for % severe malaria
-# This will be the number of cases of ...
+# Read in epi parameters 
+parameters <- read_excel("public_data/parameters.xlsx")
 
@@
 
+Short formula in R: `r round(mean(bes$pk, na.rm=TRUE),digits=2)`
+
 # Introduction
 
-The average incidence of malaria for our years of observation was `r round(mean(bes$pk, na.rm=TRUE),digits=2)`
+Mozambique is one of the 10 highest-malaria burden countries in the world, with parasite prevalence ranging from 3 to more than 50%.
+Although large drops in transmission can be achieved in a limited timeframe, indoor residual spraying and case management alone are not sufficient to attain elimination (Okumu et al, 2011).
 
-# Methodology
+Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investigação em Saúde de Manhiça (CISM).
+The main objective of this program is to assist the NMCP in adopting targeted evidence-based elimination plans through a 'learn by doing' strategy that begins in program mode and incorporates data from rigorous evaluation-into action as it is continuously being generated.
+Besides developing a fully costed, targeted, evidence-based elimination plan for Southern Mozambique, the first two years of this program have generated knowledge to inform the programmatic aspects of the elimination plan, most importantly how to interrupt transmission.
+It has also provided coordination, management and capacity building support to the NMCP, and facilitated the adoption of evidence-based malaria policies through the creation of a Malaria Technical Advisory Committee in the country.
+The  MALTEM program will provide technical, operational and financial support to design and scale up new strategies for the Mozambican National Malaria Control Program (NMCP) to achieve elimination in Maputo and significantly decrease malaria transmission rates in Gaza and Inhambane provinces by 2020.
 
-WE analyzed data from bla bla bla
+Within this context, a rigorous costing analysis needs to be carried out in order to assess the viability of MALTEM strategies and guarantee the financial sustainability of the project across time.
+Firstly, this study estimates the costs associated with malaria elimination activities in the district of Magude and models these costs, in accordance to economic theory principles, evidence from literature on malaria elimination as well as from WHO guidelines, across time.
+These costs are then compared with malaria control costs to identify potential savings of resources associated with malaria elimination, in the medium/long run.
+The whole analysis allows us assessing whether it malaria elimination (by comparing the costs of malaria elimination vs.
+costs of malaria control activities implementation and estimating the quantity and value of additional resource requirements to achieve malaria elimination), is financially affordable and sustainable over the next 10 years, but also, to which extend malaria elimination is also cost-effective compared to malaria control interventions.
 
-# Results
+#Methods
+
+##Study area and population
+
+The malaria elimination is taking place in Magude District, on the North-Western area of Maputo Province, Southern Mozambique.
+Magude District borders with the district of Massingir (Gaza Province), in the North, Manhiça District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West.
+It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds.
+The district has seven health facilities with a total of 43 beds (1.65 beds per 1,000 inhabitants) and reported 20,000 malaria cases per 100,000 population at risk in 2012.
+
+Manhiça district has been used as a comparator area for the analysis of the interventions, costing and variables of interest.
+Manhiça, located 80km north of the capital, Maputo, is a semi-rural area, with a population of 165.497 habitants.
+As of 2016, malaria remains one of the main causes of under-five morbidity and mortality in the district.
+
+##Study design
+
+The study we propose is not a randomized one.
+Malaria elimination activities will be carried out in Magude and not in a randomized way (it is not possible to randomize a district based malaria elimination campaign).
+In order to identify the impact of malaria elimination activities on indicators of impact and in order to perform the general cost-effectiveness analyses, a comparator is needed.
+There is rich data existent in Manhiça district, where CISM is located.
+However, it is subject to important noise given all biomedical research and clinical trials carried out within the district in the last 20 years.
+In consequence, we will use and ponderate data at the district level.
+
+In the estimates we will also control for potential confounders, because good data are available.
+It is possible to control, for instance, for the demographic structure of people living in the districts, daily rainfall and temperature data, etc.
+Features of the health systems of all districts in the province are also known.
+
+##Costs
+
+###Costs of malaria elimination
+
+The costing of the malaria elimination plan in Magude District consists of a detailed compilation of the estimated costs of resources used in the first 2 years and the additional resources needed to carry out the 5 year Malaria elimination program (till 2020).
+The main purpose of this costing analysis is to provide realistic information about the financial and economic resources used to execute the plan as well as the distribution of these resources across the years......
+
+###Costs of malaria case management
+
+ELISA's ref.
+Health system costs for the treatment of a malaria episode included admission costs at the maternity ward in the case of inpatients, or the cost of attendance and treatment in the case of outpatients.
+Unit costs of an admission at the maternity ward (occupied bed/day) and of an external consultation included all recurrent components and excluded capital costs.
+Recurrent costs consisted of personnel, medical, surgical and laboratory supplies among other recurrent costs.
+Unit costs were updated from the year 2000 to the year 2007 using the average annual rate inflation correction
+factor and validated during interviews with the administrative staff of the MDH [32].
+Drug costs were added to this estimate.
+Quinine was the antimalarial drug administered to admitted pregnant women and to outpatient pregnant women in the first trimester of gestation.
+The combination of SP and artesunate was administered to outpatient pregnant women who were over the first trimester of pregnancy.
+Drug costs for malaria treatment are mostly sustained by the National Health System in Mozambique, and patients only pay a small fee when attending government health facilities.
+Total admission costs were calculated by multiplying the daily unit cost of admission at the maternity ward by the average number of admission days, which were estimated through the household cost
+data collection outlined below.
+
+##Costing scenarios and sensitivity analysis
+
+Explain costs of malaria control
+
+ELISA's ref.
+The base case costing scenario relied on the following set of assumptions: a discount rate of 3 % was applied to capital costs; wastage of RDT kits, ACT treatment courses and other field work materials was assumed to be 10 %; overhead for national (and international) supervision amounted to 15 % of the total direct financial costs of the programme.
+These are believed to be conservative assumptions.
+The cost of ACT treatment courses and RDT kits
+were based on the cost, insurance and freight (cif.) price of the drug or diagnostic derived ......
+
+Elisa's ref.
+Probabilistic cost-effectiveness analysis was undertaken through Monte Carlo simulations using @Risk (version 5.0) add-in tool to Microsoft Excel (Palisade Corporation, Ithaca, NY, USA).
+The main factors affecting the cost-effectiveness of the intervention were identified in the estimated model by calculating Spearman's rank correlation coefficients as a measure of the magnitude of the association between each variable and the ICERs.
+A threshold analysis of the cost-effectiveness of IPTp-SP was performed to estimate cut-off points beyond which the
+prevention is no longer cost-effective.
+The threshold level of the ICERs used to define the intervention as cost-effective was 129 US$ per DALY averted, while 36 US$ per DALY averted was the threshold used to define the intervention as highly cost-effective.
+These threshold figures were based on previously accepted 1993 and 1996 World Bank definitions and inflated to their 2007 equivalent: 129 and 36 US$ for the year 2007 correspond to 100
+and 25 US$ for the year 1993, respectively [47-49].
+The costeffective intervention threshold was conservatively set at 100 instead of 150 US$ per DALY averted.
+Probabilistic threshold analysis was performed on SP price, other intervention costs, case fatality rate, malaria incidence, protective efficacy and antenatal clinic attendance.
+Furthermore, a one-way sensitivity analysis was undertaken on SP price.
+
+##Effectiveness
+
+All explanation on data from BES, its limitations and advantages compared to data from MALTEM's own surveillance system.......
+
+##Cost effectiveness analysis
+
+####Cost-effectiveness of MALTEM on malaria
+
+###Cost-effectiveness of MALTEM on mortality
+
+#Results
+
+##Costs
+
+###Costs elimination (baseline years)
+
+###Costs projection elimination
+
+###Costs projection control
+
+###Costs savings elimination
+
+###Effectiveness (impact)
+
+###Outcome indicators (costing outcomes)
+
+###Impact indicators
+
+###Cost-effectiveness analysis
+
+#OTHERS
 

### 2017-02-01 paper.Rmd
finished preliminary effectiveness calculations. toma epidemiologos!

@@
 author: "Laia Cirera, Joe Brew, Elisa Sicuri"
-date: "1 de fevereiro de 2017"
-output: html_document
+date: "`r Sys.Date()`"
+output:
+ tufte::tufte_html: default
+ tufte::tufte_handout:
+   citation_package: natbib
+   latex_engine: xelatex
+ tufte::tufte_book:
+   citation_package: natbib
+   latex_engine: xelatex
+bibliography: bibliography.bib
+link-citations: yes
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
-
-```{r}
 knitr::opts_chunk$set(comment = NA, 
@@
 library(tidyverse)
+library(broom)
 
@@
 
-# Read in incidence data
-bes <- read_csv("data/cleaned/cases.csv")
+# Read in incidence data and weatehr data
+source("master.R")
+bes <- df
+
+# Create variables 
+bes <- 
+  bes %>%
+  mutate(mda=ifelse(date<"2015-11-08", "pre_mda", 
+                    ifelse(date<"2016-02-04", "mda", "post_mda")),
+         intervention=district=="MAGUDE")
+
+# Create model to estimate effectiveness
+mod <-lm(p ~ mda*intervention + irs_coverage + age, data=bes, weights = population)
 ```
 
-Short formula in R: `r round(mean(bes$pk, na.rm=TRUE),digits=2)`
+Model parameters
+
+```{r}
+
+# Explore results of linear model
+# knitr::kable(tidy(summary(mod)))
+```
+
+```{r}
+
+# Use model to predict conterfactual
+other_irs <- bes %>% filter(year == 2016, district != "MAGUDE") %>% summarise(x =mean(irs_coverage, na.rm =TRUE)) %>% .$x
+counter_factual <- 
+   bes %>% filter(year == 2016, 
+                  district == "MAGUDE")
+counter_factual <- 
+  counter_factual %>%
+  mutate(irs_coverage = other_irs) %>%
+  mutate(intervention = FALSE)
+
+counter_factual$p_cf <- predict(mod, counter_factual)
+counter_factual$cases_cf <- counter_factual$p_cf*counter_factual$population 
+counter_factual$cases_averted <- counter_factual$cases_cf - counter_factual$cases
+
+# Get sum of averted cases
+ac <- counter_factual %>%
+  group_by(age)%>%
+  summarise(sum_cases_averted=sum(cases_averted))
+
+# Get the reduction in incidence
+inc <-
+  counter_factual %>%
+  group_by(age)%>%
+  summarise(reduction = (mean(p_cf) - mean(p)) / mean(p_cf) * 100)
+
+# Plot the observed and counterfactual
+plot_data <- counter_factual %>%
+  group_by(date) %>%
+  summarise(cases = sum(cases),
+            cases_cf = sum(cases_cf),
+            cases_averted = sum(cases_averted)) %>%
+  gather(key, value, cases:cases_averted) %>%
+  filter(key != "cases_averted") %>%
+  mutate(key = ifelse(key == "cases", "MALTEM", "Counter-factual"))
+
+# library(ggthemes)
+# ggplot(data=plot_data,
+#        aes(x=date,
+#            y=value,
+#            color=key))+
+#   geom_line(size =2,
+#             alpha = 0.6) + 
+#   labs(title="Observed malaria cases vs counterfactual",
+#        subtitle="Post-MDA period, 2016",
+#        x="Date",
+#        y="Malaria cases") +
+#   theme_minimal() +
+#   scale_color_manual(name = "",
+#                      values = c("red", "darkgreen"))
+
+### Create new variables in our epi data related to clinical events
+
+```
 
@@
 It has also provided coordination, management and capacity building support to the NMCP, and facilitated the adoption of evidence-based malaria policies through the creation of a Malaria Technical Advisory Committee in the country.
+
 The  MALTEM program will provide technical, operational and financial support to design and scale up new strategies for the Mozambican National Malaria Control Program (NMCP) to achieve elimination in Maputo and significantly decrease malaria transmission rates in Gaza and Inhambane provinces by 2020.
@@
 It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds.
-The district has seven health facilities with a total of 43 beds (1.65 beds per 1,000 inhabitants) and reported 20,000 malaria cases per 100,000 population at risk in 2012.
+The district has seven health facilities with a total of 43 beds ^[(1.65 beds per 1,000 inhabitants)]  and reported 20,000 malaria cases per 100,000 population at risk in 2012.
 

### 2017-02-01 get_bes_data.R

@@ -242,7 +242,8 @@ df  <- df %>%
   left_join(pop,
             by = c("year", "district", 'age'))
 
-
+# Substitute NA with 0
+df$cases[is.na(df$cases)] <- 0
 
 # Make incidence
 df <- 

### 2017-02-01 get_weather_data.R

@@ -1,172 +1,184 @@
 library(raster)
 library(R.utils)
 library(tidyverse)
+library(readr)
 
-# Define function for creating link to data
-create_url <- function(date = '2010-01-01'){
-  date <- as.Date(date)
-  url <- paste0("ftp://ftp.chg.ucsb.edu/pub/org/chg/products/CHIRPS-2.0/africa_daily/tifs/p05/",
-                format(date, '%Y'),
-                "/chirps-v2.0.",
-                format(date, '%Y'),
-                ".",
-                format(date, '%m'),
-                ".",
-                format(date, "%d"),
-                ".tif.gz")
-  return(url)
-}
-
-dates <- seq(as.Date('2010-01-01'),
-             as.Date('2016-12-31'),
-             by = 1)
-
-# If no weather_data dir, create on
-if(!dir.exists('weather_data')){
-  dir.create('weather_data')
-}
-
-for (i in 1:length(dates)){
-  this_date <- dates[i]
-  start_time <- Sys.time()
-  try({
-    # Define a file name
-    file_name <- paste0('weather_data/', this_date, '.tif')
-    # Skip if the data is already there
-    if(!file.exists(file_name)){
-      this_url <- create_url(this_date)
-      # Remove the old stuff
-      file.remove('temp.tif')
-      file.remove('temp.tif.gz')
-      # Download file
-      download.file(url = this_url,
-                    destfile = 'temp.tif.gz')
-      # Extract
-      R.utils::gunzip('temp.tif.gz')
-      # Move
-      file.copy(from = 'temp.tif',
-                to = file_name)
-      message('---------------------------------')
-      end_time <- Sys.time()
-      message(paste0('That took ',
-                     round(as.numeric(end_time - start_time), digits = 2),
-                     ' seconds.'))
+if(file.exists("weather_data/aggregated/precipitation_daily.csv") &
+   file.exists("weather_data/aggregated/precipitation_weekly.csv") &
+   file.exists("weather_data/aggregated/precipitation_monthly.csv")){
+  precipitation_daily <- read_csv("weather_data/aggregated/precipitation_daily.csv")
+  precipitation_weekly <- read_csv("weather_data/aggregated/precipitation_weekly.csv")
+  precipitation_monthly <- read_csv("weather_data/aggregated/precipitation_monthly.csv")
+  
+} else {
+  
+  # Define function for creating link to data
+  create_url <- function(date = '2010-01-01'){
+    date <- as.Date(date)
[…]

### 2017-02-01 master.R

@@ -2,7 +2,7 @@
 library(tidyverse)
 
 # BES data
-source('get_bes_data.R')
+source('get_bes_data.R', encoding = "UTF-8")
 
 # Precipitation data
 source('get_weather_data.R')
@@ -35,161 +35,161 @@ df <- df %>%
 
 # Save a copy
 write_csv(df, 'data/outputs/master.csv')
-
-# Generate a wide dataset
-if(file.exists('data/outputs/master_wide.csv')){
-  df_wide <- read_csv('data/outputs/master_wide.csv')
-} else {
-  df_wide <- df
-  
-  districts <- sort(unique(df$district))
-  years <- sort(unique(df$year))
-  
-  # Generate irs lag
-  df_wide$irs_coverage_last_year <- NA
-  for (i in 1:nrow(df)){
-    this_year <- df$year[i]
-    this_district <- df$district[i]
-    df_wide$irs_coverage_last_year[i] <-
-      df_wide$irs_coverage[df_wide$district == this_district &
-                        df_wide$year == this_year - 1][1]
-  }
-  
-  # Generate precipitation variables
-  starter <- seq(120, 1, by = -10)
-  ender <- seq(1, 120, by = 10)
-  
-  # Create empty columns
-  for (s in starter){
-    for (e in ender){
-      if(s >= e){
-        df_wide[,paste0('precipitation_from_',
-                   s, '_to_',
-                   e, '_days_ago')] <- 
-          NA
-        df_wide[,paste0('max_precipitation_from_',
-                   s, '_to_',
-                   e, '_days_ago')] <- 
-          NA
-        df_wide[,paste0('min_precipitation_from_',
-                   s, '_to_',
-                   e, '_days_ago')] <- 
-          NA
-        df_wide[,paste0('rainy_days_from_',
-                   s, '_to_',
-                   e, '_days_ago')] <- 
-          NA
-        df_wide[,paste0('dry_days_from_',
-                   s, '_to_',
-                   e, '_days_ago')] <- 
-          NA
-      }
-    }
-  }
-  
-  # Populate emtpy columns
-  for (s in starter){
-    for (e in ender){
-      message(paste0(s, '  ', e))
-      if(s >= e){
-        for (i in 1:nrow(df_wide)){
-          this_district <- df_wide$district[i]
-          this_date <- df_wide$date[i]
-          
-          the_data <-
-            precipitation_daily %>%
-            filter(date >= (this_date - s),
-                   date <= (this_date - e),
[…]

### 2017-02-01 paper.Rmd
more modifications with laia, cost savings identified

@@
 library(broom)
+library(ggthemes)
 
@@
 ```{r}
-
 # Use model to predict conterfactual
@@
                   district == "MAGUDE")
+results <- list()
+for (i in 2016:2025){
+  new_year <- counter_factual %>%
+    mutate(year = i) 
+  results[[i]] <- new_year
+}
+counter_factual <- bind_rows(results)
+counter_factual$date <- 
+  as.Date(paste0(counter_factual$year, substr(counter_factual$date, 5, 10)))
+
 counter_factual <- 
@@
 ac <- counter_factual %>%
+  filter(year == 2016) %>%
   group_by(age)%>%
@@
   counter_factual %>%
+  filter(year == 2016) %>%
   group_by(age)%>%
@@
 
-# Plot the observed and counterfactual
-plot_data <- counter_factual %>%
-  group_by(date) %>%
-  summarise(cases = sum(cases),
-            cases_cf = sum(cases_cf),
-            cases_averted = sum(cases_averted)) %>%
-  gather(key, value, cases:cases_averted) %>%
-  filter(key != "cases_averted") %>%
-  mutate(key = ifelse(key == "cases", "MALTEM", "Counter-factual"))
+### Create new variables in our epi data related to clinical events
+counter_factual <- 
+  counter_factual %>%
+  mutate(cases_non_severe = cases * parameters$value[parameters$parameter == 'p_non_sev'],
+         cases_non_severe_cf = cases_cf * parameters$value[parameters$parameter == 'p_non_sev'],
+         cases_severe = cases * parameters$value[parameters$parameter == 'p_sev'],
+         cases_severe_cf = cases_cf * parameters$value[parameters$parameter == 'p_sev'],
+         deaths = cases * parameters$value[parameters$parameter == 'cfr'],
+         deaths_cf = cases_cf * parameters$value[parameters$parameter == 'cfr'],
+         cost_outpatient = cases_non_severe * parameters$value[parameters$parameter == 'cost_outp_usd'],
+         cost_outpatient_cf = cases_non_severe_cf * parameters$value[parameters$parameter == 'cost_outp_usd'],
+        cost_inpatient = cases_severe * parameters$value[parameters$parameter == 'cost_inp_usd'],
+        cost_inpatient_cf = cases_severe_cf * parameters$value[parameters$parameter == 'cost_inp_usd']) %>%
+  mutate(cost_clinical = cost_inpatient + cost_outpatient,
+         cost_clinical_cf = cost_inpatient_cf + cost_outpatient_cf)
 
-# library(ggthemes)
-# ggplot(data=plot_data,
-#        aes(x=date,
-#            y=value,
-#            color=key))+
-#   geom_line(size =2,
-#             alpha = 0.6) + 
-#   labs(title="Observed malaria cases vs counterfactual",
-#        subtitle="Post-MDA period, 2016",
-#        x="Date",
-#        y="Malaria cases") +
-#   theme_minimal() +
-#   scale_color_manual(name = "",
-#                      values = c("red", "darkgreen"))
+# Dollars per case averted
+# We have cases_averted
+# We also have clinical costs saved
+# We need program costs
 
-### Create new variables in our epi data related to clinical events
+# 2016
+# dpca = program costs + real clinical costs / cf program costs + cf clinical costs
 
+# We need to differentiate between program costs (MALTEM) and intervention costs
+intervention_costs <- costs[!costs$elimination_activity %in% c('Entomology', 'REACT'),]
+
+intervention_costs <-
+  intervention_costs %>%
+  # MALTEM COSTS
+  mutate(cost_maltem = 
+              (resource_use * 
+              q_maltem *
+              p_maltem *
+              yearly_regularity / 
+              useful_life)) %>%
+  mutate(transport_meticais = ifelse(transport_cost, 0.25 * cost_maltem, 0),
+         import_meticais = ifelse(import_tax, 0.4 * cost_maltem, 0),
+         operating_meticais = ifelse(operating_cost, 0.1 * cost_maltem, 0),
+         maintenance_meticais = ifelse(maintenance_cost, 0.15 * cost_maltem, 0)) %>%
+  mutate(cost_maltem = 
+           cost_maltem +
+           transport_meticais +
+           import_meticais +
+           operating_meticais +
+           maintenance_meticais) %>%
+  mutate(cost_basic = 
+           (resource_use * 
+              q_basic *
+              (p_basic_perc * p_maltem) *
+              yearly_regularity / 
+              useful_life)) %>%
+    mutate(transport_meticais_basic = ifelse(transport_cost, 0.25 * cost_maltem, 0),
+         import_meticais_basic = ifelse(import_tax, 0.2 * cost_maltem, 0),
+         operating_meticais_basic = ifelse(operating_cost, 0.1 * cost_maltem, 0),
+         maintenance_meticais_basic = ifelse(maintenance_cost, 0.15 * cost_maltem, 0)) %>%
+    mutate(cost_basic = 
+           cost_basic +
+           transport_meticais_basic +
+           import_meticais_basic +
+           operating_meticais_basic +
+           maintenance_meticais_basic) %>%
+  dplyr::select(elimination_activity,
+                budget_category,
+                baseline_year,
+                item,
+                description,
+                cost_basic,
+                cost_maltem) %>%
+  mutate(cost_basic = ifelse(is.na(cost_basic), 0, cost_basic))
+
+# Get total intervention costs
+total_intervention_cost <- 
+  intervention_costs[,c('cost_basic', 'cost_maltem')]
+total_intervention_cost <-
+  tidy(apply(total_intervention_cost, 2, sum, na.rm = TRUE) / 75)
+
+# Get a counterfactula for government
+cf_costs <- intervention_costs %>%
+  filter(elimination_activity %in% c('IRS', 'LLIN')) %>%
+  summarise(x = sum(cost_basic) / 75)
+cf_costs <- data_frame(names = 'counter_factual',
+                       x = cf_costs$x)
+
+total_intervention_cost <-
+  bind_rows(total_intervention_cost,
+            cf_costs)
+names(total_intervention_cost)[2] <- 'cost'
+
+ggplot(data = total_intervention_cost,
+       aes(x= names,
+           y = cost)) +
+  geom_bar(stat = 'identity') +
+  labs(x = 'Scenario',
+       y = 'US Dollars',
+       title = 'Comparative costs of malaria control/elimination approaches',
+       subtitle = 'USD') +
+  geom_label(aes(label = round(cost)))
+
+# Add clinical costs to our total_intervention_cost dataset
+clinical_costs <- 
+  counter_factual %>%
+  filter(year == 2016) %>%
+  summarise(counter_factual = sum(counter_factual$cost_clinical_cf),
+            cost_basic = sum(counter_factual$cost_clinical),
+            cost_maltem = sum(counter_factual$cost_clinical))
+clinical_costs <- tidy(t(clinical_costs))
+names(clinical_costs) <- c('names', 'clinical_cost')
+total_intervention_cost <-
+  left_join(x = total_intervention_cost,
+            y = clinical_costs,
+            by = 'names')
+
+# Get total cost (program plus clinic)
+total_intervention_cost$total_cost <-
+  total_intervention_cost$cost +
+  total_intervention_cost$clinical_cost
+
+# Get total cases for 2016 under both scenarios
+x <- counter_factual %>%
+  filter(year == 2016) %>%
+  summarise(cases_maltem = sum(cases),
+            cases_cf = sum(cases_cf))
+
+DT::datatable(total_intervention_cost)
 ```
@@
 
-Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investigação em Saúde de Manhiça (CISM).
+Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investiga??o em Sa?de de Manhi?a (CISM).
 The main objective of this program is to assist the NMCP in adopting targeted evidence-based elimination plans through a 'learn by doing' strategy that begins in program mode and incorporates data from rigorous evaluation-into action as it is continuously being generated.
@@
 The malaria elimination is taking place in Magude District, on the North-Western area of Maputo Province, Southern Mozambique.
-Magude District borders with the district of Massingir (Gaza Province), in the North, Manhiça District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West.
+Magude District borders with the district of Massingir (Gaza Province), in the North, Manhi?a District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West.
 It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds.
@@
 
-Manhiça district has been used as a comparator area for the analysis of the interventions, costing and variables of interest.
-Manhiça, located 80km north of the capital, Maputo, is a semi-rural area, with a population of 165.497 habitants.
+Manhi?a district has been used as a comparator area for the analysis of the interventions, costing and variables of interest.
+Manhi?a, located 80km north of the capital, Maputo, is a semi-rural area, with a population of 165.497 habitants.
 As of 2016, malaria remains one of the main causes of under-five morbidity and mortality in the district.
@@
 In order to identify the impact of malaria elimination activities on indicators of impact and in order to perform the general cost-effectiveness analyses, a comparator is needed.
-There is rich data existent in Manhiça district, where CISM is located.
+There is rich data existent in Manhi?a district, where CISM is located.
 However, it is subject to important noise given all biomedical research and clinical trials carried out within the district in the last 20 years.
@@
 Elisa's ref.
-Probabilistic cost-effectiveness analysis was undertaken through Monte Carlo simulations using @Risk (version 5.0) add-in tool to Microsoft Excel (Palisade Corporation, Ithaca, NY, USA).
+Probabilistic cost-effectiveness analysis was undertaken through Monte Carlo simulations using @Risk (version 5.0) add-in tool to Microsoft Excel.
+(Palisade Corporation, Ithaca, NY, USA).
 The main factors affecting the cost-effectiveness of the intervention were identified in the estimated model by calculating Spearman's rank correlation coefficients as a measure of the magnitude of the association between each variable and the ICERs.
@@
 All explanation on data from BES, its limitations and advantages compared to data from MALTEM's own surveillance system.......
+
+```{r}
+# Plot the observed and counterfactual
+plot_data <- counter_factual %>%
+  group_by(date) %>%
+  summarise(cases = sum(cases),
+            cases_cf = sum(cases_cf),
+            cases_averted = sum(cases_averted)) %>%
+  gather(key, value, cases:cases_averted) %>%
+  filter(key != "cases_averted") %>%
+  mutate(key = ifelse(key == "cases", "MALTEM", "Counter-factual"))
+
+ggplot(data=plot_data,
+       aes(x=date,
+           y=value,
+           color=key))+
+  geom_line(size =2,
+            alpha = 0.6) +
+  labs(title="Observed malaria cases vs counterfactual",
+       subtitle="Post-MDA period, 2016",
+       x="Date",
+       y="Malaria cases") +
+  theme_minimal() +
+  scale_color_manual(name = "",
+values = c("red", "darkgreen"))
+```
 
@@
 
-#OTHERS
-
-Our results show that the effectiveness of MALTEM's interventions is as follows:
-
-```{r}
-plot(x=costs$year , y=costs$`Case fatality rate`, type="l",
-     xlab="YEAR", ylab="rate", main="Case fatality rate ACROSS TIME",
-     col="blue")
-```
-
 And here is another figure.

### 2017-03-07 create_risk_map.R
risk map animation

@@ -1,0 +1,145 @@
+# Prepare main data
+source('master.R')
+
+# Get locations
+coords <- read_csv('public_data/coordinates_of_maputo_districts.csv')
+
+# Join to data
+df <- left_join(x = df,
+                y = coords,
+                by = 'district')
+
+# Rename
+df$lng <- df$x
+df$lat <- df$y
+
+# Get map of Maputo province
+library(cism)
+library(automap)
+library(maptools)
+map <- moz1
+map <- map[map@data$NAME_1 == 'Maputo',]
+map_fortified <- broom::tidy(map, region = 'NAME_1')
+
+
+# Define function for risk map
+risk_map <- function(data,
+                     map = map){
+  
+  # Group by location and get overall
+  grouped <- data %>%
+    group_by(lng, lat) %>%
+    summarise(cases = sum(cases),
+              population = sum(population)) %>%
+    ungroup %>%
+    mutate(pk = cases / population * 1000)
+  
+  # Perform kriging
+  # Krieging
+  sp_df <- grouped
+  coordinates(sp_df) <- ~lng + lat
+  proj4string(sp_df) <- proj4string(map)
+  sp_df <- sp_df[!is.na(sp_df$pk),]
+  sp_points <- SpatialPoints(sp_df)
+  
+  # Create a fine grid
+  pixels_per_side = 150
+  bottom.left = bbox(map)[,1]  #apply(sp_points@coords,2,min)
+  top.right = bbox(map)[,2] #apply(sp_points@coords,2,max)
+  margin = abs((top.right-bottom.left))/10
+  bottom.left = bottom.left-margin
+  top.right = top.right+margin
+  pixel.size = abs(top.right-bottom.left)/pixels_per_side
+  g = GridTopology(cellcentre.offset=bottom.left,
+                   cellsize=pixel.size,
+                   cells.dim=c(pixels_per_side,pixels_per_side))
+  
+  grid_points = SpatialPoints(g)
+  proj4string(grid_points) <- proj4string(map)
+  in_points = !is.na(over(grid_points,polygons(map)))
+  fit_points = SpatialPoints(as.data.frame(grid_points)[in_points,])
+  proj4string(fit_points) <- proj4string(map)
+  
+  # make projected before krieging
+  new_pro <- CRS("+init=epsg:3036")
+  proj4string(fit_points) <-
+    proj4string(map) <- 
+    proj4string(sp_df) <- new_pro
+  
[…]

### 2017-03-09 paper.Rmd
more work on weather

@@
                     ifelse(date<"2016-02-04", "mda", "post_mda")),
-         intervention=district=="MAGUDE")
+         intervention=district=="MAGUDE",
+         magude = ifelse(district == 'MAGUDE', 'Magude', 'Not Magude'))
 
 # Create model to estimate effectiveness
-mod <-lm(p ~ mda*intervention + irs_coverage + age, data=bes, weights = population)
+mod <-lm(pk ~ mda*intervention + irs_coverage, data=bes, weights = population)
 ```
@@
 results <- list()
-for (i in 2016:2025){
+for (i in 2010:2020){
   new_year <- counter_factual %>%
@@
 ```{r}
+x <- counter_factual
+x$predicted <- predict(mod, x)
+long <- gather(x, key, value, cases:predicted)
+long$value <- as.numeric(long$value)
+long <- long %>%
+  mutate(value = ifelse(key == 'pk', NA, value))
+
+y <- bes
+long2 <- gather(y, key, value, cases:predicted)
+long2$value <- as.numeric(long2$value)
+
+z <- rbind(long %>%
+             filter(year >= 2017),
+           long2 %>%
+             filter(year <= 2016))
+
+ggplot(data = z %>%
+         filter(key %in% c('predicted', 'pk'),
+                district == 'MAGUDE'),
+       aes(x = date,
+           y = value,
+           group = key,
+           color = key)) +
+  geom_line() +
+  facet_wrap(~age)
+```
+
+```{r}
 # Plot the observed and counterfactual
@@
 
+```{r}
+x <- bes %>%
+    filter(district != 'NAMAACHA') %>%
+  # filter(district != 'MANHICA') %>%
+  group_by(date,
+           # year, month,
+           # age, 
+           magude) %>%
+  summarise(#pk = weighted.mean(pk, w = population),
+            cases = sum(cases),
+            pop = sum(population),
+            precip = mean(precip)) %>%
+  ungroup %>%
+  mutate(p = cases / pop * 1000) #%>%
+  # mutate(date = as.Date(paste0(year, '-',
+  #                              '06',
+  #                              # month, 
+  #                              '-01')))# %>%
+  # filter(date >= '2014-07-01')
+
+library(ggthemes)
+library(cism)
+ggplot(data = x,
+       aes(x = date,
+           color = magude,
+           y = p)) +
+  geom_line(alpha = 0.6) +
+  # geom_smooth() +
+  # facet_wrap(~age, ncol = 1) +
+  # theme_bw()  +
+  scale_color_manual(name = '',
+                     values = c('blue', 'red')) +
+  # theme_fivethirtyeight() +
+  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-12-31'))),
+             lty = 2, alpha = 0.2) +
+  geom_vline(xintercept = as.numeric(as.Date(paste0('2015-03-15'))),
+             color = 'darkgreen') +
+  labs(x = 'Week',
+       y = 'Weekly incidence per 1,000',
+       title = 'Weekly clinical malaria',
+       subtitle = 'Province of Maputo, all ages, weekly') +
+  geom_label(data = data_frame(x = as.Date(paste0('2015-03-15')),
+                               y = 15,
+                               label = 'MDA'),
+             aes(x = x,
+                 y = y,
+                 label = label),
+             color = 'black') +
+  theme_cism()
+```
 # Discussion
+
+```{r}
+bes <- bes %>% filter(district != 'NAMAACHA')
+# Prediction
+bes$predicted <- predict(mod, bes)
+
+x <- bes %>%
+    filter(district != 'NAMAACHA') %>%
+  # filter(district != 'MANHICA') %>%
+  group_by(date,
+           # year, month,
+           # age, 
+           magude) %>%
+  summarise(#pk = weighted.mean(pk, w = population),
+            cases = sum(cases),
+            pop = sum(population),
+            precip = mean(precip),
+            precip_risk = mean(precipitation_risk),
+            predicted = mean(predicted)) %>%
+  ungroup %>%
+  mutate(p = cases / pop * 1000)
+#%>%
+  # mutate(date = as.Date(paste0(year, '-',
+  #                              '06',
+  #                              # month, 
+  #                              '-01')))# %>%
+  # filter(date >= '2014-07-01')
+
+library(ggthemes)
+ggplot(data = x,
+       aes(x = date,
+           color = magude,
+           y = predicted)) +
+  geom_line(alpha = 0.6) +
+  # geom_smooth() +
+  # facet_wrap(~age, ncol = 1) +
+  # theme_bw()  +
+  scale_color_manual(name = '',
+                     values = c('blue', 'red')) +
+  theme_fivethirtyeight() +
+  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-12-31'))),
+             lty = 2, alpha = 0.6) +
+  geom_vline(xintercept = as.numeric(as.Date(paste0('2015-03-15')))) +
+  geom_line(data = x,
+            aes(x = date,
+                y = p,
+                color = magude),
+                lty = 2,
+            alpha = 0.5)
+
+```
+
+```{r}
+# Precipitation
+x <- bes %>%
+  group_by(year, month) %>%
+  summarise(p = mean(precipitation)) %>%
+  mutate(date = as.Date(paste0(year, '-', month, '-01')))
+ggplot(data = x,
+       aes(x = date,
+           y = p)) +
+  geom_bar(stat = 'identity') 
+```

### 2017-03-09 create_risk_map.R

@@ -21,10 +21,18 @@ map <- moz1
 map <- map[map@data$NAME_1 == 'Maputo',]
 map_fortified <- broom::tidy(map, region = 'NAME_1')
 
+map2 <- moz2
+map2 <- map2[map2@data$NAME_1 == 'Maputo',]
+map2_fortified <- broom::tidy(map2, region = 'NAME_2')
 
 # Define function for risk map
 risk_map <- function(data,
                      map = map){
+  
+  # Jitter the data
+  data <- data %>%
+    mutate(lng = jitter(lng, factor = 50),
+           lat = jitter(lat, factor = 50))
   
   # Group by location and get overall
   grouped <- data %>%
@@ -89,9 +97,9 @@ risk_map <- function(data,
     geom_tile(aes(fill=pred),color=NA) +
     # stat_contour(aes(z=pred), bins=nbin, color="#999999") +
     scale_fill_gradient2(name = 'Hotspot\nindex',
-                         midpoint=mean(interp_data$pred),
-                         # limits = c(-1,5),
-                         # midpoint = 2,
+                         # midpoint=mean(interp_data$pred),
+                         limits = c(-1,7),
+                         midpoint = 2,
                          low="white",
                          mid="yellow",
                          high="red") +
@@ -134,12 +142,32 @@ for (i in 1:length(dates)){
     filter(date == dates[i])
   r <- risk_map(data = data,
                 map = map) +
-    labs(title = dates[i])
+    labs(title = format(dates[i], '%B %d, %Y')) +
+    geom_polygon(data = map2_fortified,
+                 aes(x = long,
+                     y = lat,
+                     group = group),
+                 color = 'black',
+                 fill = NA)
   r
   ggsave(paste0('risk_map_animation/', dates[i], '.png'))
   # assign(paste0('g', i),
          # r)
 }
+
+# Static map
+data <- df %>%
+  filter(date >= '2016-12-15')
+r <- risk_map(data = data,
+              map = map) +
+  labs(title = 'December 2016') +
+  geom_polygon(data = map2_fortified,
+               aes(x = long,
+                   y = lat,
+                   group = group),
+               color = 'black',
+               fill = NA)
+r
 # ggplot_list <- Filter(function(x) is(x, "ggplot"), mget(ls()))
 
 # Run command line: convert -delay 10 -loop 0 *.png result.gif

### 2017-03-09 get_weather_data.R

@@ -348,7 +348,8 @@ if(file.exists("weather_data/aggregated/precipitation_daily.csv") &
     precipitation_weekly <- 
       precipitation_daily %>%
       group_by(year, week, district) %>%
-      summarise(precipitation = sum(precipitation, na.rm = TRUE),
+      summarise(date = first(date),
+                precipitation = sum(precipitation, na.rm = TRUE),
                 precipitation_risk = length(which(precipitation_risk)) / n(),
                 avg_temp_max = mean(temp_max, na.rm = TRUE),
                 max_temp = max(temp_max, na.rm = TRUE),

### 2017-03-09 master.R

@@ -10,12 +10,6 @@ source('get_weather_data.R')
 # IRS data
 source('get_irs_data.R')
 
-# Join malaria cases with precipitation and weather
-df <-
-  left_join(x = df,
-            y = precipitation_weekly,
-            by = c('district', 'year', 'week'))
-
 # Join with IRS
 df <- 
   left_join(x = df,
@@ -35,111 +29,7 @@ df <- df %>%
 
 # Save a copy
 write_csv(df, 'data/outputs/master.csv')
-# 
-# # Generate a wide dataset
-# if(file.exists('data/outputs/master_wide.csv')){
-#   df_wide <- read_csv('data/outputs/master_wide.csv')
-# } else {
-#   df_wide <- df
-#   
-#   districts <- sort(unique(df$district))
-#   years <- sort(unique(df$year))
-#   
-#   # Generate irs lag
-#   df_wide$irs_coverage_last_year <- NA
-#   for (i in 1:nrow(df)){
-#     this_year <- df$year[i]
-#     this_district <- df$district[i]
-#     df_wide$irs_coverage_last_year[i] <-
-#       df_wide$irs_coverage[df_wide$district == this_district &
-#                         df_wide$year == this_year - 1][1]
-#   }
-#   
-#   # Generate precipitation variables
-#   starter <- seq(120, 1, by = -10)
-#   ender <- seq(1, 120, by = 10)
-#   
-#   # Create empty columns
-#   for (s in starter){
-#     for (e in ender){
-#       if(s >= e){
-#         df_wide[,paste0('precipitation_from_',
-#                    s, '_to_',
-#                    e, '_days_ago')] <- 
-#           NA
-#         df_wide[,paste0('max_precipitation_from_',
-#                    s, '_to_',
-#                    e, '_days_ago')] <- 
-#           NA
-#         df_wide[,paste0('min_precipitation_from_',
-#                    s, '_to_',
-#                    e, '_days_ago')] <- 
-#           NA
-#         df_wide[,paste0('rainy_days_from_',
-#                    s, '_to_',
-#                    e, '_days_ago')] <- 
-#           NA
-#         df_wide[,paste0('dry_days_from_',
-#                    s, '_to_',
-#                    e, '_days_ago')] <- 
-#           NA
-#       }
-#     }
-#   }
-#   
-#   # Populate emtpy columns
-#   for (s in starter){
-#     for (e in ender){
-#       message(paste0(s, '  ', e))
[…]

### 2017-03-09 weather_model.R

@@ -1,0 +1,163 @@
+# BES data
+source('get_bes_data.R', encoding = "UTF-8")
+
+# Aggregate by age
+df <- df %>%
+  group_by(date, district) %>%
+  summarise(year = first(year),
+            month = first(month),
+            week = first(week),
+            day= first(day),
+            cases = sum(cases),
+            population = sum(population)) %>%
+  mutate(p = cases / population,
+         pk = cases / population * 1000)
+
+# Precipitation data
+source('get_weather_data.R')
+
+# Get locations
+coords <- read_csv('public_data/coordinates_of_maputo_districts.csv')
+
+# Join to data
+df <- left_join(x = df,
+                y = coords,
+                by = 'district')
+
+# Rename
+df$lng <- df$x
+df$lat <- df$y
+
+# Generate a wide weather dataset
+if('wide_weather.csv' %in% dir()){
+  wide_weather <- read_csv('weather_data/aggregated/wide_weather.csv')
+} else {
+  
+  districts <- sort(unique(precipitation_weekly$district))
+  years <- sort(unique(precipitation_weekly$year))
+  
+  # Generate precipitation variables
+  starter <- seq(120, 0, by = -5)
+  ender <- seq(0, 120, by = 5)
+  
+  counter <- 0
+  total <- length(starter) * length(ender)
+  # Create empty columns
+  for (s in starter){
+    for (e in ender){
+      if(s >= e){
+        precipitation_weekly[,paste0('precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('max_precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('min_precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('rainy_days_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('dry_days_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('min_temp_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        precipitation_weekly[,paste0('max_temp_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
[…]

### 2017-03-10 paper.Rmd
major weather overhaul

@@
 
-ggplot(data = bes,
+ggplot(data = df,
        aes(x = date,
@@
 ```{r}
-x <- bes %>%
+x <- df %>%
+  mutate(magude = ifelse(district == 'MAGUDE', 'Magude', 'Not Magude')) %>%
     filter(district != 'NAMAACHA') %>%

### 2017-03-10 get_bes_data.R

@@ -1,6 +1,9 @@
 library(tidyverse)
 library(readxl)
 
+# Source helpers
+source('helpers.R')
+
 # Read in modulo basico data
 # x <- read.csv('data/modulo_basico/Compiled_Dados Provincia Maputo.csv', skip = 2)
 mb <- read_excel('data/modulo_basico/Compiled_Dados Provincia Maputo.xlsx',
@@ -259,12 +262,23 @@ df$day <- as.numeric(format(df$date, '%d'))
 df <- df %>%
   filter(date <= '2016-12-31')
 
+# Get a date helper for standardizing weekly dates to saturday
+date_helper <- create_date_helper()
+
+df <- df %>% ungroup
+
+df <- df %>%
+  dplyr::select(-date) %>%
+  left_join(date_helper,
+            by = c('year', 'week'))
+
 # Write data
 write_csv(df, 'data/outputs/cases.csv')
 
 # Remove unecessary objects
-rm(d2, mb, pop,
-   final_list, i, sh,
+rm(d2, date_helper, mb, pop,
+   final_list, i, sh, counter,
+   this_dow,
    read_population)
 
 # # See incidence

### 2017-03-10 get_weather_data.R

@@ -1,403 +1,206 @@
-library(raster)
-library(R.utils)
+# Packages
 library(tidyverse)
-library(readr)
 
-if(file.exists("weather_data/aggregated/precipitation_daily.csv") &
-   file.exists("weather_data/aggregated/precipitation_weekly.csv") &
-   file.exists("weather_data/aggregated/precipitation_monthly.csv")){
-  precipitation_daily <- read_csv("weather_data/aggregated/precipitation_daily.csv")
-  precipitation_weekly <- read_csv("weather_data/aggregated/precipitation_weekly.csv")
-  precipitation_monthly <- read_csv("weather_data/aggregated/precipitation_monthly.csv")
-  
+# Source helpers
+source('helpers.R')
+
+# If the data has already been processed, simply read it in
+if('mozambican_weather.csv' %in% dir('noaa_data')){
+  weather <- read_csv('noaa_data/mozambican_weather.csv')
 } else {
   
-  # Define function for creating link to data
-  create_url <- function(date = '2010-01-01'){
-    date <- as.Date(date)
-    url <- paste0("ftp://ftp.chg.ucsb.edu/pub/org/chg/products/CHIRPS-2.0/africa_daily/tifs/p05/",
-                  format(date, '%Y'),
-                  "/chirps-v2.0.",
-                  format(date, '%Y'),
-                  ".",
-                  format(date, '%m'),
-                  ".",
-                  format(date, "%d"),
-                  ".tif.gz")
-    return(url)
-  }
-  
-  dates <- seq(as.Date('2010-01-01'),
-               as.Date('2016-12-31'),
-               by = 1)
-  
-  # If no weather_data dir, create on
-  if(!dir.exists('weather_data')){
-    dir.create('weather_data')
-  }
-  
-  for (i in 1:length(dates)){
-    this_date <- dates[i]
-    start_time <- Sys.time()
-    try({
-      # Define a file name
-      file_name <- paste0('weather_data/', this_date, '.tif')
-      # Skip if the data is already there
-      if(!file.exists(file_name)){
-        this_url <- create_url(this_date)
-        # Remove the old stuff
-        file.remove('temp.tif')
-        file.remove('temp.tif.gz')
-        # Download file
-        download.file(url = this_url,
-                      destfile = 'temp.tif.gz')
-        # Extract
-        R.utils::gunzip('temp.tif.gz')
-        # Move
-        file.copy(from = 'temp.tif',
-                  to = file_name)
[…]

### 2017-03-10 get_wide_weather.R

@@ -1,0 +1,310 @@
+# This should be run after "get_weather_data.R"
+# It assumes object in memory from that
+source('get_weather_data.R')
+
+library(tidyverse)
+
+# Generate a wide weather dataset
+if('mozambican_weather_wide.csv' %in% dir('noaa_data/')){
+  wide_weather <- read_csv('noaa_data/mozambican_weather_wide.csv')
+} else {
+  
+  # Trim down weekly weather 
+  weather_weekly <- 
+    weather_weekly %>%
+    filter(!is.na(date)) %>%
+    filter(date >= '2010-01-01')
+  
+  # Generate time ranges
+  starter <- seq(120, 5, by = -5)
+  ender <- seq(5, 120, by = 5)
+  
+  counter <- 0
+  combos <- expand.grid(s = starter,
+                        e = ender)
+  combos <- combos %>% filter(s >= e)
+  total <- nrow(combos)
+
+  # Create empty columns
+  for (s in starter){
+    for (e in ender){
+      if(s >= e){
+        weather_weekly[,paste0('precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('max_precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('min_precipitation_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('rainy_days_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('dry_days_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('min_temp_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('max_temp_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('avg_temp_from_',
+                                     s, '_to_',
+                                     e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('avg_dew_point_from_',
+                               s, '_to_',
+                               e, '_days_ago')] <-
+          NA
+        weather_weekly[,paste0('avg_wind_speed_from_',
+                               s, '_to_',
[…]

### 2017-03-10 helpers.R

@@ -1,0 +1,118 @@
+# Define function for creating date helper
+create_date_helper <- 
+  function(){
+    message('Standardizing dates so that all weekly data use ',
+            'the date of the Saturday in that week (weeks are Sunday-Saturday')
+    # Standardize dates (use dates ending on Saturday of each week)
+    date_helper <- 
+      data_frame(date = seq(as.Date('2009-01-01'),
+                            as.Date('2017-12-31'),
+                            by = 1))
+    date_helper$dow <- weekdays(date_helper$date)
+    date_helper <-
+      date_helper %>%
+      mutate(year = as.numeric(format(date, '%Y')),
+             week = as.numeric(format(date, '%U')))
+    
+    date_helper <- date_helper %>%
+      group_by(year, week) %>%
+      mutate(has_saturday = 'Saturday' %in% dow) %>%
+      filter(has_saturday) %>%
+      ungroup %>%
+      dplyr::select(-has_saturday)
+    
+    date_helper$new_date <- as.Date(NA)
+    for (i in 1:nrow(date_helper)){
+      counter <- i
+      this_dow <- date_helper$dow[counter]
+      while(this_dow != 'Saturday'){
+        counter <- counter + 1
+        this_dow <- date_helper$dow[counter]
+      }
+      date_helper$new_date[i] <- as.Date(date_helper$date[counter])
+    }
+    date_helper <-
+      date_helper %>%
+      dplyr::select(new_date, year, week) %>%
+      rename(date = new_date)
+    date_helper <- date_helper[!duplicated(date_helper$date),]
+    return(date_helper)
+  }
+
+
+# Define function to replace NAs (coded in noaa as 99.9 by NOAA)
+detect_noaa_na <- function(x){
+  x <- as.character(x)
+  y <- gsub('.', '', x, fixed = TRUE)
+  oks <- y == x
+  out <- unlist(lapply(strsplit(y, ''), function(x){all(x == '9')}))
+  out[oks] <- FALSE
+  out
+}
+
+# Define functions for converting from farenheit to celcius
+f_to_c <- function(x){
+  x <- x - 32
+  x <- x * (5/9)
+  x
+}
+
+# Define function for converting from inches to milimeters
+i_to_m <- function(x){
+  x <- x * 25.4
+  x
+}
+
[…]

### 2017-03-10 master.R

@@ -10,6 +10,11 @@ source('get_weather_data.R')
 # IRS data
 source('get_irs_data.R')
 
+df <-		
+  left_join(x = df,		
+            y = weather,		
+            by = c('district', 'date'))
+
 # Join with IRS
 df <- 
   left_join(x = df,

### 2017-03-10 weather_functions.R

@@ -1,0 +1,415 @@
+##########
+# This script has the code for models 
+
+
+##########
+# Random Forest 
+##########
+rfPred <- function(model_data,
+                   nfolds,
+                   cutoff,
+                   iterations) 
+{
+  
+  model <- list()
+  best_features <- list()
+  importance <- list()
+  train.mse <- list()
+  test.mse <- list()
+  train.predictions <- list()
+  test.predictions <- list()
+  train.ground_truth <- list()
+  test.ground_truth <- list()
+  test_acc <- list()
+  test_stats  <- list()
+  
+  
+  # Remove some useless variables
+  model_data <-
+    model_data %>%
+    dplyr::select(-year, -month,
+                  -week, -day,
+                  -p, -pk,
+                  -x, -y,
+                  -lng, -lat)
+  
+  # remove all rows with NAs - you can also impute
+  model_data <- model_data[complete.cases(model_data),]
+  
+  # Make sure it's a dataframe
+  model_data <- data.frame(model_data)
+  
+  ## get features - remove outcome and date (because it has too many factor levels for the models)
+  feature_index <- !grepl('cases|date|district', colnames(model_data))
+  selected_features <- colnames(model_data)[feature_index]
+  
+  for (i in 1:iterations) {
+    
+    set.seed(i)
+    
+    train_index <- sample(nrow(model_data), nrow(model_data) *cutoff, replace = F)
+    
+    # set numer of folds and parameters for training data
+    fitControl <- trainControl( 
+      method = "repeatedcv",  # could train on boostrap resample, here use repeated cross validation.
+      number = nfolds,      
+      repeats = 1,
+      allowParallel = TRUE
+    )
+    
+    y <- model_data$cases[train_index] # gets training outcome
+    
+    # mtry: Number of variables randomly sampled as candidates at each split.
+    # ntree: Number of trees to grow - these are standard settings for RF
+    mtry <- sqrt(ncol(model_data[train_index, selected_features]))
+    tunegrid <- expand.grid(.mtry=mtry)
+    
+    model[[i]] <- train(x = model_data[train_index, selected_features]
+                        , y = y
+                        , method = "rf"
+                        , trControl = fitControl
+                        , tuneGrid = tunegrid
+                        , importance = T
+                        , verbose = FALSE)
+    
+    temp <- varImp(model[[i]])[[1]]
[…]

### 2017-03-10 weather_model.R

@@ -1,163 +1,0 @@
-# BES data
-source('get_bes_data.R', encoding = "UTF-8")
-
-# Aggregate by age
-df <- df %>%
-  group_by(date, district) %>%
-  summarise(year = first(year),
-            month = first(month),
-            week = first(week),
-            day= first(day),
-            cases = sum(cases),
-            population = sum(population)) %>%
-  mutate(p = cases / population,
-         pk = cases / population * 1000)
-
-# Precipitation data
-source('get_weather_data.R')
-
-# Get locations
-coords <- read_csv('public_data/coordinates_of_maputo_districts.csv')
-
-# Join to data
-df <- left_join(x = df,
-                y = coords,
-                by = 'district')
-
-# Rename
-df$lng <- df$x
-df$lat <- df$y
-
-# Generate a wide weather dataset
-if('wide_weather.csv' %in% dir()){
-  wide_weather <- read_csv('weather_data/aggregated/wide_weather.csv')
-} else {
-  
-  districts <- sort(unique(precipitation_weekly$district))
-  years <- sort(unique(precipitation_weekly$year))
-  
-  # Generate precipitation variables
-  starter <- seq(120, 0, by = -5)
-  ender <- seq(0, 120, by = 5)
-  
-  counter <- 0
-  total <- length(starter) * length(ender)
-  # Create empty columns
-  for (s in starter){
-    for (e in ender){
-      if(s >= e){
-        precipitation_weekly[,paste0('precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('max_precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('min_precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('rainy_days_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('dry_days_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('min_temp_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        precipitation_weekly[,paste0('max_temp_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
[…]

### 2017-05-15 compare_with_opd.R
previous state

@@ -105,3 +105,111 @@ g1
 ggsave('~/Desktop/magude.png')
 g2
 ggsave('~/Desktop/manhica.png')
+
+# GET incidence in ij
+x <- 
+  opd %>%
+  filter(!is.na(date)) %>%
+  mutate(date >= as.Date('2010-01-01')) %>%
+  mutate(year_month = lendable::date_truncate(date, 'month')) %>%
+  mutate(month = as.numeric(format(date, '%m'))) %>%
+  filter(place == 'Ilha Josinha') %>%
+  mutate(age_group = ifelse(age < 12, '0-0.99',
+                            ifelse(age < 60, '1-4.99',
+                                   ifelse(age >= 60, '5+', NA)))) %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  filter(year >= 2010) %>%
+  group_by(date, age_group) %>%
+  summarise(visits = n(),
+            malaria = length(which(malaria))) %>%
+  ungroup %>%
+  mutate(p = malaria / visits * 100)
+
+ggplot(data = x %>% filter(age_group != '5+'),
+       aes(x = date,
+           y = malaria,
+           group = age_group,
+           color = age_group)) +
+  geom_line(alpha = 0.6) +
+  # geom_point() +
+  theme_cism() +
+  labs(x = 'Age group',
+       y = 'Daily malaria cases')
+ggsave('~/Desktop/daily.png')
+
+write_csv(x, '~/Desktop/daily.csv')
+
+# GET incidence in ij
+x <- 
+  opd %>%
+  filter(!is.na(date)) %>%
+  mutate(date >= as.Date('2010-01-01')) %>%
+  mutate(year_month = lendable::date_truncate(date, 'month')) %>%
+  mutate(month = as.numeric(format(date, '%m'))) %>%
+  filter(place == 'Ilha Josinha') %>%
+  mutate(age_group = ifelse(age < 12, '0-0.99',
+                            ifelse(age < 60, '1-4.99',
+                                   ifelse(age >= 60, '5+', NA)))) %>%
+  mutate(year = as.numeric(format(date, '%Y'))) %>%
+  group_by(year, year_month, month, age_group) %>%
+  summarise(visits = n(),
+            malaria = length(which(malaria))) %>%
+  ungroup %>%
+  mutate(p = malaria / visits * 100) %>%
+  filter(year >= 2010)
+
+ggplot(data = x, #%>% filter(age_group != '5+'),
+       aes(x = year_month,
[…]

### 2017-05-15 master.R

@@ -99,6 +99,112 @@ df_wide <-
 # Write wide weather too
 write_csv(df_wide, 'data/outputs/cases_population_weather_wide_itn_irs.csv')
 
+
+# # Make a plot of Magude over time
+# x <- df %>%
+#   filter(district == 'MAGUDE') %>%
+#   group_by(date, disease) %>% 
+#   summarise(cases = sum(cases),
+#             population = sum(population)) %>%
+#   ungroup %>%
+#   mutate(k = cases / population * 1000)
+# 
+# library(cism)
+# ggplot(data = x %>%
+#          filter(disease == 'MALARIA'),
+#        aes(x = date,
+#            y = k)) +
+#   # geom_point() +
+#   geom_line(color = 'darkorange',
+#             alpha = 0.8) +
+#   theme_cism() +
+#   labs(x = 'Date',
+#        y = 'Weekly incidence per 1,000',
+#        title = 'Malaria cases',
+#        subtitle = 'Magude, Mozambique')
+# 
+# 
+# # Currency
+# owd <- getwd()
+# setwd('../lendable/app')
+# library(lendable)
+# mzn <- fetch_db(query = "SELECT * FROM currency WHERE currency_type ='MZN'")
+# setwd(owd)
+# # Get to value of mzn
+# mzn$mzn <- 1/ mzn$usd 
+# mzn <- mzn %>%
+#   filter(date >= as.Date('2010-01-01'),
+#          mzn <= 0.15)
+# ggplot(data = mzn,
+#        aes(x = date,
+#            y = mzn)) +
+#   geom_line(color = 'darkorange',
+#             alpha = 0.8) +
+#   labs(x = 'Date',
+#        y = '?',
+#        title = 'Huge reduction',
+#        subtitle = 'Thanks, MALTEM!') +
+#   theme_cism() 
+# 
+# # Map of gaza and maputo
+# gm <- cism::moz1_fortified %>%
+#   filter(id %in% c('Maputo', 'Gaza'))
+# ggplot() +
+#   geom_polygon(data = gm,
+#                aes(x = long,
+#                    y = lat,
+#                    group = group,
+#                    fill = id),
+#                alpha = 0.7) +
+#   geom_polygon(data = mag3_fortified,
+#                aes(x = long,
+#                    y = lat,
+#                    group = group),
+#                fill = 'red') +
+#   theme_cism_map() +
+#   coord_map() +
+#   scale_fill_manual(name = '',
[…]

### 2017-05-15 get_irs_data.R
fixing duplicates issue

@@ -2,150 +2,171 @@ library(tidyverse)
 library(readr)
 library(readxl)
 
-irs <- read_csv('data/vector_control_ministry/IRS_mozambique_clean_final.csv')
-
-# Fix Matola
-irs$district[irs$district == 'M.Matola'] <- 'Matola'
-irs$province[irs$district == 'Matola'] <- 'Maputo'
-
-# Keep only gaza and maputo
-irs <- irs %>%
-  filter(province %in% c('Gaza', 'Maputo'))
-
-# Standardize names to match those in bes
-irs$district <- toupper(irs$district)
-# irs$province <- toupper(irs$province)
-irs_districts <- sort(unique(irs$district))
-
-# Clean dates
-irs$start_date <- irs$end_date <- NA
-irs$start_date <- as.Date(paste0(irs$start_year, 
-                         '-',
-                         irs$start_month,
-                         '-',
-                         irs$start_day))
-irs$end_date <- as.Date(paste0(irs$end_year, 
-                         '-',
-                         irs$end_month,
-                         '-',
-                         irs$end_day))
-
-irs$end_date[is.na(irs$end_date)] <-
-  irs$start_date[is.na(irs$end_date)] +
-  irs$number_of_days_activity[is.na(irs$end_date)]
-
-irs$start_date[is.na(irs$start_date)] <-
-  irs$end_date[is.na(irs$start_date)] -
-  irs$number_of_days_activity[is.na(irs$start_date)]
-
-# For those days with no start/end date, estimate
-irs$no_irs_date <- is.na(irs$start_date)
-
-# The average is 3 months, and it's usually at end of year
-# so we're guessing Oct 1 through Dec 31
-irs$start_date <- as.character(irs$start_date)
-irs$end_date <- as.character(irs$end_date)
-irs$start_date <-
-  ifelse(is.na(irs$start_date),
-         paste0(irs$start_year, '-10-01'),
-         as.character(irs$start_date))
-irs$end_date <-
-  ifelse(is.na(irs$end_date),
-         paste0(irs$start_year, '-12-31'),
-         as.character(irs$end_date))
-irs$start_date <- as.Date(irs$start_date)
-irs$end_date <- as.Date(irs$end_date)
-
-# If the end date is before the start date, 
-# take 3 months after the start date
-irs$end_date[irs$start_date > irs$end_date] <-
-  irs$start_date[irs$start_date > irs$end_date] + 90
-
[…]

### 2017-05-15 get_bes_data.R
updates

@@ -535,6 +535,16 @@ bes <- bes %>%
                 cases) %>%
   arrange(date, province, district, age_group)
 
+# Remove the one weird dup for MANDLAKAZI
+bes <- bes %>%
+  arrange(date) %>%
+  mutate(dummy = 1) %>%
+  group_by(date, district, age_group, disease) %>%
+  mutate(dummy = cumsum(dummy)) %>%
+  filter(dummy == 1) %>%
+  ungroup %>%
+  dplyr::select(-dummy)
+
 # Join population to data
 df  <- bes %>%
   left_join(pop %>% dplyr::select(-province),
@@ -545,6 +555,8 @@ df <-
   df %>%
   mutate(p = cases / population) %>%
   mutate(pk = p * 1000)
+
+
 
 # Remove unecessary objects
 rm(district_matcher,

### 2017-05-15 get_irs_data.R

@@ -190,7 +190,10 @@ if('irs_data_cleaned.RData' %in% dir('data/binaries')){
       }
     }
   }
+  
+  # Join to dh
+  irs <- irs %>%
+    left_join(dh)
   save(irs,
        file = 'data/binaries/irs_data_cleaned.RData')
 }
-irs$irs_no_date <- NULL

### 2017-05-15 master.R

@@ -79,7 +79,14 @@ df <-
             y = weather_weekly,		
             by = c('district', 'date'))
 
+# Make an older and younger dataset
+df_young <-df %>% filter(age_group == '0-4')
+df_old <- df %>% filter(age_group == '5+')
+
 # Write data for sharing with collaborators
+write_csv(df_young, 'data/outputs/cases_population_weather_itn_irs_young_only.csv')
+write_csv(df_old, 'data/outputs/cases_population_weather_itn_irs_old_only.csv')
+
 write_csv(df, 'data/outputs/cases_population_weather_itn_irs.csv')
 write_csv(bes, 'data/outputs/cases_only.csv')
 write_csv(pop, 'data/outputs/population.csv')

### 2017-05-16 reports/weather_prediction/weather_prediction.Rmd
done

---
title: "Weather accurately predicts clinical malaria in Magude"
author: "MALTEM"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

# Packages 
library(databrew)
library(cism)
library(dplyr)
library(ggplot2)
library(ggmap)
library(gsheet)
library(readr)

# Options for this document (ie, whether to show code, errors, etc.)
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE,
               eval = TRUE,
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)

# Read in wide weather data
source('../../weather_functions.R', chdir = TRUE)
source('../../master.R', chdir = TRUE)

# Get weather for Magude
if('mag_weather_wide.RData' %in% dir()){
  load('mag_weather_wide.RData')
} else {
  mag_weather <- cism::get_weather(station = 'FQMA',
                                 start_year = 2010,
                                 end_year = 2017)

  # Rename columns
  mag_weather <-
    mag_weather %>%
    dplyr::select(date, 
                  temp_max,
                  temp_mean,
                  temp_min,
                  precipitation) %>%
    rename(temp = temp_mean)
  # # Keep only Saturdays
  # mag_weather <-
  #   mag_weather %>%
  #   filter(weekdays(date) == 'Saturday') 
  # Add district
  mag_weather$district <- 'MAGUDE'

  # Make mag wide
  mag_weather_wide <- make_weather_wide(weather = mag_weather,
                                        remove_before = '2010-01-01',
                                        go_forward = TRUE)
  save(mag_weather_wide,
       file = 'mag_weather_wide.RData')
}

# Keep only saturdays
mag_weather_wide <-
  mag_weather_wide %>%
  filter(weekdays(date) == 'Saturday')

# Model the outcome
model_data <-
  mag_weather_wide %>% left_join(
      df %>%
  filter(disease == 'MALARIA',
         district == 'MAGUDE') %>%
  group_by(date, province, district) %>%
  summarise(cases = sum(cases)) %>%
  ungroup 
  )

model_data <-
  model_data[,grepl('cases|precip|temp|date', names(model_data))]
model_data$mda <- model_data$date >= '2015-10-01'
old_model_data <- model_data
model_data <- model_data %>% filter(date <= '2016-09-01')
model_data <- model_data[complete.cases(model_data),]
dates <- model_data$date; model_data$date <- NULL
model_data$temp <- model_data$temp_max <-
  model_data$temp_min <- model_data$precipitation <- NULL

library(randomForest)
fit <- randomForest(cases ~ ., data = model_data)
model_data <- old_model_data
model_data$predicted <- predict(fit, newdata = model_data)
model_data$predicted[model_data$predicted <0] <- 0
long <- model_data %>%
  dplyr::select(date, predicted, cases) %>%
  gather(key,
         value,
         predicted:cases)

```

# Executive summary

A machine learning model (using a "random forest" algorithm), trained on weather data (precipitation, temperature variation) and MDA data (pre/during MALTEM dummy) can predict the number of clinical malaria cases in Magude with a fairly high degree of accuracy.

# Data sources

## Weather

We use weather from the Maputo airport weather station, obtained through the `get_data` function in the publicly available `cism` package in R.

## Surveillance

We use BES data from the Mozambican Ministry of Health. We have this data on record through the end of 2016.

# Feature generation

Our weather data contain only 4 variables of interest: precipitation, mean temperature, maximum temperature, and minimum temperature. However, we generate lagged and aggregated features for periods of interest. These include the following (hopefully self-explanatory) variables:

```{r}
sort(unique(names(mag_weather_wide)[7:ncol(mag_weather_wide)]))
```

We intentionally blind the model to (a) any historical case data, and (b) any weather data in the previous 20 days. Thanks to the latter, we are able to predict 20 days in advance (since the model requires only data greater than 20 days old).

# Methods

We run a random forest regression with all features. Random forest has an advantage over linear regression because (a) it better handles thresholds, (b) it better handles interactions.

# Results

The below plot shows the number of true cases (per BES data) and predicted cases (per the algorithm) over the last 7 years.

```{r}
g <- ggplot(data = long,
       aes(x = date,
           y = value,
           color = key)) +
  geom_line(alpha = 0.6) +
  scale_color_manual(name = '',
                     values = c('red', 'blue')) +
  theme_cism() +
  labs(x = 'Date',
       y = 'Weekly cases',
       title = 'Clinical malaria in Magude',
       subtitle = 'Observed vs. predicted cases')
g
```

The below plot is identical to the above, but zoomed in just to the last 9 months. Note that the algorithm correctly predicted both (a) low malaria incidence during the summer, and (b) the recent rise in cases. The model fails at the low end (ie, it often predicts 0 when in reality there are always several cases), but appears to catch peaks appropriately.

Also note that this model is _completely_ blind to data after 2016-09-01. Even wihtout any epi information, it predicts a startling rise in malaria cases in Magude during the month of April and May, which we have observed in reality.

```{r}
g + 
  xlim(as.Date(c('2016-10-01', '2017-05-16'))) +
  geom_point(alpha = 0.6) +
  labs(title = 'Zoom-in: Clinical malaria in Magude') +
  ylim(0, 250)

```

The below table shows the the number of predicted and observed cases by week.

```{r}
DT::datatable(model_data %>%
                dplyr::select(date, cases, predicted) %>%
                arrange(desc(date)) %>%
                filter(!is.na(predicted)) %>%
                mutate(predicted = round(predicted)))
```

The below chart shows the predicted and observed cases for each year separately.

```{r}
model_data <-
  model_data %>%
  mutate(day_number = as.numeric(format(date, '%j'))) %>%
  mutate(year = as.numeric(format(date, '%Y'))) %>%
  mutate(fake_date = as.Date(paste0('2016-01-01')) + day_number) %>%
  mutate(blind = ifelse(date <= '2016-09-01', 'in sample', 'out of sample'))
long <- 
  model_data %>%
  dplyr::select(date, 
                year, 
                fake_date,
                predicted,
                cases) %>%
  gather(key,
         value,
         predicted:cases) %>%
  mutate(year2 = ifelse(year == 2017, '2017', 'other years'))
ggplot(data = long,
       aes(x = fake_date,
           y = value,
           color = key)) +
  geom_line() +
  facet_wrap(~year) +
    scale_color_manual(name = '',
                     values = c('red', 'blue')) +
  theme_cism() +
  labs(x = 'Date',
       y = 'Weekly cases',
       title = 'Clinical malaria in Magude',
       subtitle = 'Observed vs. predicted cases') +
  theme(axis.text.x = element_text(angle = 90))
```

The below chart reflects model accuracy / stability. The model performs fairly well across both low and high incidence periods. The color of each point reflects whether this period was in the training sample or not (ie, we are more concerned about "out of sample" performance than in sample). Note that this does not include any data from 2017.

```{r}
ggplot(data =model_data,
       aes(x = predicted,
           y = cases,
           color = blind)) +
  geom_point(alpha = 0.6) +
  theme_cism() +
  labs(x = 'Predicted',
       y = 'Observed',
       title= 'Model performance: all time',
       subtitle = 'Magude clinical malaria cases') +
  scale_color_manual(name = '',
                     values = c('darkorange', 'darkgreen'))
```

# Discussion

## Take-aways

There is a lot more work to do on this. But this "minimal viable product" goes to show that:

- Weather data is fairly predictive of malaria incidence.
- Weather data is easily usable.
- Weather data can be used not only to _describe_, but also to _predict_.
- Had we implemented a forecasting algorithm earlier, we could have predicted the recent spike in Malaria cases in mid-March (even though incidence was low then). 
- The model correctly predicted the late April / early May spike.
- The model predicts that the incidence of malaria will subside slightly in late May and early June, but will remain well above February/April levels.

## Limitations

- The model is still showing weakness in the intervention period (it over-predicts systematically).
- The model appears to have incorrectly predicted a January 2017 spike; this may be correctable by better incorporation of MDA timing (currently using only a dummy).
- No concept of confidence-bounding has yet been introduced.

### 2017-05-16 get_weather_data.R

@@ -4,6 +4,7 @@ library(sp)
 
 # Source helpers
 source('helpers.R')
+source('weather_functions.R')
 
 # If the data has already been processed, simply read it in
 if('mozambican_weather.csv' %in% dir('noaa_data')){

### 2017-05-16 get_wide_weather.R

@@ -4,174 +4,18 @@
 
 library(tidyverse)
 
+
+
 # Generate a wide weather dataset
 if('mozambican_weather_wide.csv' %in% dir('noaa_data/')){
   wide_weather <- read_csv('noaa_data/mozambican_weather_wide.csv')
 } else {
-  
-  # Trim down weekly weather 
-  weather_weekly <- 
-    weather_weekly %>%
-    filter(!is.na(date)) %>%
-    filter(date >= '2010-01-01')
-  
-  # Generate time ranges
-  starter <- seq(100, 0, by = -10)
-  ender <- seq(0, 100, by = 10)
-  
-  counter <- 0
-  combos <- expand.grid(s = starter,
-                        e = ender)
-  combos <- combos %>% filter(s >= e)
-  total <- nrow(combos)
-
-  # Create empty columns
-  for (s in starter){
-    for (e in ender){
-      if(s >= e){
-        weather_weekly[,paste0('precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('max_precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('min_precipitation_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('rainy_days_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('dry_days_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('min_temp_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('max_temp_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('avg_temp_from_',
-                                     s, '_to_',
-                                     e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('avg_dew_point_from_',
-                               s, '_to_',
-                               e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('avg_wind_speed_from_',
-                               s, '_to_',
-                               e, '_days_ago')] <-
-          NA
-        weather_weekly[,paste0('max_wind_speed_from_',
[…]

### 2017-05-16 master.R

@@ -7,8 +7,8 @@ source('get_bes_data.R', encoding = "UTF-8")
 # Weather data
 source('get_weather_data.R')
 
-# Lagged weather data
-source('get_wide_weather.R')
+# # Lagged weather data
+# source('get_wide_weather.R')
 
 # IRS data
 source('get_irs_data.R')
@@ -93,18 +93,18 @@ write_csv(pop, 'data/outputs/population.csv')
 # Write csv for weather data
 write_csv(weather_weekly, 'data/outputs/weather_weekly.csv')
 write_csv(weather, 'data/outputs/weather_daily.csv')
-write_csv(wide_weather, 'data/outputs/weather_wide.csv')
+# write_csv(wide_weather, 'data/outputs/weather_wide.csv')
 # Write csv for itn data
 write_csv(itn, 'data/outputs/itn.csv')
 # Write csv for irs data
 write_csv(irs, 'data/outputs/irs.csv')
 
-# Join with wide weather
-df_wide <- 
-  left_join(x = df,
-            y = wide_weather)
-# Write wide weather too
-write_csv(df_wide, 'data/outputs/cases_population_weather_wide_itn_irs.csv')
+# # Join with wide weather
+# df_wide <- 
+#   left_join(x = df,
+#             y = wide_weather)
+# # Write wide weather too
+# write_csv(df_wide, 'data/outputs/cases_population_weather_wide_itn_irs.csv')
 
 
 # # Make a plot of Magude over time

### 2017-05-16 weather_functions.R

@@ -401,7 +401,196 @@ plotModel <- function(result_list,
   abline(0,1, lty = 3)
   corr <- round(cor(unlist(result_list[[4]]), unlist(result_list[[6]])), 2)
   legend("topleft", legend = paste0('correlation = ', corr), cex = 1, bty = 'n')
-  
-  
 }
 
+########
+# Define function for getting wide weather
+make_weather_wide <- function(weather,
+                              remove_before = '2010-01-01',
+                              go_forward = FALSE){
+  
+  # Generate time ranges
+  lowest <- 20
+  highest <- 100
+  starter <- seq(highest, lowest, by = lowest * -1)
+  ender <- seq(lowest, highest, by = lowest)
+  
+  # Add rows to weather
+  if(go_forward){
+    new_row <- weather[1,]
+    max_date <- max(weather$date)
+    new_row[,c('temp_max','temp','temp_min',
+               'precipitation')] <- NA
+    new_rows <- list()
+    for(i in 1:(lowest*2)){
+      this_row <- new_row
+      this_row$date <- max_date + i
+      new_rows[[i]] <- this_row
+    }
+    new_rows <- bind_rows(new_rows)
+    weather <- bind_rows(weather,
+                         new_rows)
+  }
+  
+  
+  counter <- 0
+  combos <- expand.grid(s = starter,
+                        e = ender)
+  combos <- combos %>% filter(s > e)
+  total <- nrow(combos)
+  
+  # Create empty columns
+  for (s in starter){
+    for (e in ender){
+      if(s > e){
+        weather[,paste0('precipitation_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('max_precipitation_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('min_precipitation_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('rainy_days_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('dry_days_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('min_temp_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
+        weather[,paste0('max_temp_from_',
+                        s, '_to_',
+                        e, '_days_ago')] <-
+          NA
[…]

### 2017-07-16 reports/misc_analysis/misc_analysis.Rmd
questions for london

---
title: "MALTEM effectiveness analysis"
author: "Brew, Cirera, Sicuri, Thomas"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

# Packages 
library(databrew)
library(cism)
library(dplyr)
library(ggplot2)
library(ggmap)
library(gsheet)
library(readr)

# Options for this document (ie, whether to show code, errors, etc.)
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE,
               eval = TRUE,
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
# Read in wide weather data
# source('../../weather_functions.R', chdir = TRUE)
source('../../master.R', chdir = TRUE)

# Define a theme
theme_maltem <- ggthemes::theme_hc
```

# Visual overview

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_line(alpha = 0.8,
            color = 'darkred') +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  facet_wrap(~district) +
  theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, age_group, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = age_group,
           color = age_group)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('darkorange', 'darkgreen')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~district) +
  # theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem()

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place, age_group) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~age_group)
```

# When did the decline begin in Magude?

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. other districts',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  group_by(date) %>%
  summarise(magude_to_other = p[place == 'Magude'] / 
              p[place == 'Other districts'] * 100)

g <- ggplot(data = x,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  geom_hline(yintercept = 100,
             col = 'darkred',
             alpha = 0.6,
             lty = 2) +
  labs(x = 'Date',
       y = 'Percent') +
  labs(title = 'Monthly incidence: Magude as % of other districts',
       subtitle = 'All age groups') +
  theme_maltem()

g + xlim(as.Date('2013-01-01'), max(x$date))

```

# Magude vs. Manhiça

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(district %in% c('MAGUDE', 'MANHICA')) %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  
                        ifelse(district == 'MANHICA', 'Manhiça', NA))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Monthly incidence: Magude vs. Manhiça',
       subtitle = 'All age groups, zoom-in') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

```

# Weather

## Precipitation

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

g <- ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = precipitation,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = precipitation,
           group = district)) +
  labs(x = 'Date',
       y = 'Milimeters',
       title = 'Precipitation: Magude vs. other districts',
       subtitle = 'Monthly average precipitation, Magude in red') +
  theme_maltem()
g 
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')

y <- x %>%
  group_by(date) %>%
  summarise(magude_to_other = mean(precipitation[district == 'MAGUDE']) / 
              mean(precipitation[district != 'MAGUDE']) * 100)
g <- 
  ggplot(data = y,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  theme_maltem() +
  labs(x = 'Date',
       y = 'Percent',
       title = 'Average monthly precipitation: Magude vs other',
       subtitle = 'Magude as percentage of monthly mean') +
  geom_hline(yintercept = 100,
             color = 'darkred',
             lty = 2,
             alpha = 0.7) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3)
g
g + 
  xlim(as.Date('2014-01-01'),
               max(x$date)) +
  labs(title = 'Zoom-in of above chart')
```

## Temperature

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = temp,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = temp,
           group = district)) +
  labs(x = 'Date',
       y = 'Degrees (celcius)',
       title = 'Average temperature: Magude vs. other districts',
       subtitle = 'Monthly average, Magude in red') +
  theme_maltem() +
  xlim(as.Date('2014-01-01'),
               max(x$date))
```

## Is seasonality constant by year?

```{r}
x <- 
  df %>%
  group_by(year, month) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(year = factor(year))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = month,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  geom_point() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  labs(x = 'Month',
       y = 'Average incidence',
       title = 'Incidence by month, all districts',
       subtitle = 'Wide variability in monthly incidence') +
  theme_maltem()
```

## Does 4 week long-and-lagged precipitation predict incidence?

Yes, but not in the way we'd expect. When we adjust for the fixed effect of month, we get a negative correlation between lagged precipitation and incidence.

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  filter(date <= '2015-06-01') %>%
  mutate(year_month = as.Date(paste0(year, '-', month, '-01'))) %>%
  arrange(date) %>%
  group_by(district) %>%
  mutate(lag4 = lag(precipitation, n = 4, default = NA),
            lag5 = lag(precipitation, n = 5, default = NA),
            lag6 = lag(precipitation, n = 6, default = NA),
            lag7 = lag(precipitation, n = 7, default = NA)) %>%
  mutate(p = sum(cases) / sum(population) * 1000) %>%
  mutate(n_nas = ifelse(is.na(lag4), 1, 0) + 
           ifelse(is.na(lag5), 1, 0) + 
           ifelse(is.na(lag6), 1, 0) + 
           ifelse(is.na(lag7), 1, 0)) %>%
  mutate(n_ok = 4 - n_nas) %>%
  mutate(precipitation_lagged = ifelse(is.na(lag4), 0, lag4) +
           ifelse(is.na(lag5), 0, lag5) +
           ifelse(is.na(lag6), 0, lag6) +
           ifelse(is.na(lag7), 0, lag7) / 
           n_ok) %>%
  filter(!is.na(precipitation_lagged),
         !is.na(p))

fit <- lm(p ~ precipitation_lagged + 
            factor(month), data = x)
broom::tidy(fit)
```

# Proximity to South Africa

# Questions

1. What are the mathematics of the method by which we assign weights to synthetic control districts? Are we concerned that so few districts get any weight?

### 2017-07-19 paper.Rmd
update

@@
 author: "Laia Cirera, Joe Brew, Elisa Sicuri"
-date: "`r Sys.Date()`"
+date: "`r March 31, 2017`"
 output:
@@
   bes %>%
-  mutate(mda=ifelse(date<"2015-11-08", "pre_mda", 
-                    ifelse(date<"2016-02-04", "mda", "post_mda")),
+  mutate(mda=ifelse(date<"2015-08-01", "pre_maltem", 'maltem'),
          intervention=district=="MAGUDE",
@@
 
+# Create a time since 2010 variable
+bes <- bes %>%
+  mutate(time_since_2010 = as.numeric(date - as.Date('2010-01-01')))
+
+# Make a season variable
+bes <- bes %>%
+  mutate(season = ifelse(month %in% 1:5, 'high', 'low'))
+
 # Create model to estimate effectiveness
-mod <-lm(pk ~ mda*intervention + irs_coverage, data=bes, weights = population)
+mod <-lm(pk ~ mda*intervention + 
+           irs_coverage + 
+           season +
+           factor(month) +
+           time_since_2010, 
+         data=bes %>%
+           filter(date >= '2012-01-01'), 
+         weights = population)
 ```
@@
 counter_factual$p_cf <- predict(mod, counter_factual)
-counter_factual$cases_cf <- counter_factual$p_cf*counter_factual$population 
+counter_factual$cases_cf <-(counter_factual$p_cf/1000)*counter_factual$population 
 counter_factual$cases_averted <- counter_factual$cases_cf - counter_factual$cases
@@
 DT::datatable(total_intervention_cost)
+
+# Plot of the counterfactual and real
+real <- bes %>%
+  filter(district == 'MAGUDE') %>%
+  mutate(status = 'observed')
+cf <- counter_factual %>%
+  # filter(year == 2016) %>%
+  filter(year < 2017) %>%
+  mutate(status = 'counterfactual') %>%
+  mutate(cases = cases_cf,
+         p = p_cf)
+plot_data <- bind_rows(real, cf)
+
+x <- plot_data %>%
+  filter(date >= '2015-08-01') %>%
+  group_by(status, year) %>%
+  summarise(total_cases = sum(cases))
+
+days_of_observation <- as.numeric(as.Date('2016-12-31') - as.Date('2015-08-01'))
+
+plot_data <- plot_data %>%
+  group_by(year, week, status) %>%
+  summarise(month = first(month),
+            date = first(date),
+            cases = sum(cases),
+            population = sum(population)) %>%
+  ungroup %>%
+  mutate(p = cases / population) %>%
+  mutate(pk = p * 1000)
+
+library(cism)
+plot_data <- plot_data %>%
+ filter(date >= '2014-07-01')
+g <- ggplot(data = plot_data ,
+       aes(x = date,
+           y = pk,
+           group = status,
+           color = status)) +
+  geom_line(alpha = 0.6) +
+  theme_cism() +
+  labs(x = 'Date',
+       y = 'Weekly incidence per 1000',
+       title = 'Observed vs. counterfactual',
+       subtitle = 'Preliminary effectivenesss model') +
+  scale_color_manual(name = 'Status',
+                     values = c('darkorange', 'darkgreen'))
+g +
+  geom_vline(xintercept = c(as.numeric(as.Date('2015-08-01')),
+                            as.numeric(as.Date('2015-11-01'))),
+             alpha = 0.5,
+             lty = 2)
+
+# By mes
+pdm <- plot_data %>%
+  mutate(date = as.Date(paste0(year, '-',
+                               month, '-01'))) %>%
+  group_by(date, status) %>%
+  summarise(pk = mean(pk))
+ggplot(data = pdm ,
+       aes(x = date,
+           y = pk,
+           group = status,
+           color = status)) +
+  geom_line(alpha = 0.6) +
+  theme_cism() +
+  labs(x = 'Date',
+       y = 'Monthly incidence per 1,000',
+       title = 'Preliminary effectiveness model',
+       subtitle = 'Magude district') +
+  scale_color_manual(name = '',
+                     values = c('darkorange', 'darkgreen')) +
+    geom_vline(xintercept = c(as.numeric(as.Date('2015-08-01')),
+                            as.numeric(as.Date('2015-11-09')),
+                            as.numeric(as.Date('2016-12-02'))),
+             alpha = 0.5,
+             lty = 2)
+
+# Get difference b/w 2016 counterfactual and real
+x <- counter_factual %>%
+  filter(year == 2016) %>%
+  summarise(cf = sum(cases_cf, na.rm = TRUE),
+            cr = sum(cases)) %>%
+  mutate(cases_averted = cf - cr)
+
+# write a csv for laia
+write_csv(plot_data, '~/Desktop/cf_vs_observed.csv')
 ```

### 2017-07-19 reports/misc_analysis/misc_analysis.Rmd
update

@@
             factor(month), data = x)
-broom::tidy(fit)
+DT::datatable(broom::tidy(fit))
 ```
@@
 
+We calculate the distance from the centroid of each district to the South African border.
+
+```{r}
+# Interpolate for our locations
+library(sp)
+x <- cism::moz2
+x <- x[x@data$NAME_1 %in% c('Maputo', 'Gaza'),]
+coords <- data.frame(coordinates(x))
+names(coords) <- c('x', 'y')
+coords$district <- x@data$NAME_2
+
+# Standardize names
+coords$district <- toupper(coords$district)
+coords$district[coords$district == 'CHÓKWÈ'] <- 'CHOKWE'
+coords$district[coords$district == 'GUIJÁ'] <- 'GUIJA'
+coords$district[coords$district == 'XAI-XAI'] <- 'XAI-XAI DISTRICT'
+coords$district[coords$district == 'MANHIÇA'] <- 'MANHICA'
+coords$district[coords$district == 'MATUTUÍNE'] <- 'MATUTUINE'
+coords$district[coords$district == 'MANDLAKAZI'] <- 'MANJACAZE'
+
+# Add a row for matola
+matola <- data_frame(x = 32.46580,
+                     y = -25.92391,
+                     district = 'MATOLA')
+coords <- bind_rows(coords, 
+                    matola)
+# Add a row for xai-xai city
+xxc <- coords %>% filter(district == 'XAI-XAI DISTRICT') %>%
+  mutate(district = 'XAI-XAI CITY')
+coords <- bind_rows(coords,
+                    xxc)
+# Make spatial
+coords <- coords %>%
+  mutate(lng = x,
+         lat = y)
+coordinates(coords) <- ~x+y
+proj4string(coords) <- proj4string(moz3)
+
+# Get South Africa
+sa <- cism::africa
+sa <- sa[sa@data$COUNTRY == 'South Africa',]
+
+# Get projected versions
+sa_proj <- spTransform(sa,
+                    CRS("+init=epsg:3347"))
+coords_proj <- spTransform(coords,
+                    CRS("+init=epsg:3347"))
+
+# Get distance to south Africa
+coords@data$distance_to_south_africa <- NA
+for (i in 1:nrow(coords)){
+  this_distance <- rgeos::gDistance(coords_proj[i,], sa_proj)
+  coords@data$distance_to_south_africa[i] <- this_distance / 10000
+}
+
+afr <- cism::africa
+afr <- afr[afr@data$COUNTRY %in% c('South Africa', 'Mozambique'),]
+plot(afr, col = adjustcolor(c('blue', 'green'), alpha.f = 0.3))
+points(coords, col = 'red',
+       pch = '.',
+       cex = 2)
+```
+
+The distances can be seen in the below table.
+
+```{r}
+DT::datatable(coords@data %>% dplyr::select(district, distance_to_south_africa) %>%
+                mutate(distance_to_south_africa = round(distance_to_south_africa, digits = 2)) %>%
+  rename(District = district,
+         KM = distance_to_south_africa)) 
+```
+
+Visually, we assess the hypothesis that the decline over time is steeper for those districts bordering South Africa through a series of charts.
+
+```{r}
+x <- df %>%
+  filter(disease == 'MALARIA') %>%
+  left_join(coords@data) %>%
+  group_by(district, year, distance_to_south_africa) %>%
+  summarise(p = sum(cases) / sum(population) * 1000)
+
+ggplot(data = x,
+       aes(x = distance_to_south_africa,
+           y = p)) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(n = 5, span = 10, se = FALSE,
+              color = 'darkred',
+              alpha = 0.5,
+              lty = 1) +
+  facet_wrap(~year) +
+  theme_maltem() +
+  labs(x = 'Kilometers to South Africa',
+       y = 'Weekly incidence per K',
+       title = 'Protective effect of South African proximity',
+       subtitle = 'One point per district per year')
+
+```
+
+As an alternative to the above, we can visualize the same data with the same aesthetics, but overlaying each year.
+
+```{r}
+library(RColorBrewer)
+cols <- colorRampPalette(brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
+ggplot(data = x,
+       aes(x = distance_to_south_africa,
+           y = p,
+           color = factor(year))) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(n = 5, span = 10, se = FALSE,
+              alpha = 0.5,
+              lty = 1) +
+  theme_maltem() +
+  labs(x = 'Kilometers to South Africa',
+       y = 'Weekly incidence per K',
+       title = 'Protective effect of South African proximity',
+       subtitle = 'One point per district per year') +
+  scale_color_manual(name = 'Year',
+                     values = cols)
+```
+
+Could the proximity to South Africa partly explain the pre-MALTEM decline in incidence in Magude?
+We'll examine this by modeling incidence as a linear function of district, distance, the interaction of year and distance, and month, in effect constructing a "counter-factual" for Magude that takes into account the secular decline in incidence associated with proximity to South Africa.
+For the model's purpose, we take into account only those 4 years from 2012-06-01 through 2016-06-01 (so as to not include the effect of MALTEM).
+Our model formula predicts incidence as a function of the interactions of season/year and distance/year, with the fixed effect of month.
+
+```{r}
+model_data <- 
+  df %>%
+  filter(disease == 'MALARIA') %>%
+  left_join(coords@data) %>%
+  mutate(season = ifelse(month %in% 1:4, 'rainy', 'not-rainy')) %>%
+  mutate(month = factor(month)) %>%
+  group_by(date, year, 
+           season,
+           month,
+           district, distance_to_south_africa) %>%
+  summarise(p = sum(cases) / sum(population) * 1000) %>%
+  ungroup
+
+fit <- lm(p ~ year*log(distance_to_south_africa)  + year + season*year + month, 
+          data = model_data %>%
+            filter(date <= '2016-06-01',
+                   date >= '2012-06-01'))
+
+DT::datatable(broom::tidy(fit))
+
+model_data$predicted <- predict(fit, model_data)
+
+ggplot() +
+  geom_line(data = model_data %>%
+         filter(district == 'MAGUDE'),
+       aes(x = date,
+           y = predicted),
+       alpha = 0.8) +
+  geom_line(data = model_data %>%
+         filter(district == 'MAGUDE'),
+       aes(x = date,
+           y = p),
+       color = 'red',
+       alpha = 0.8) +
+  labs(title = 'Weekly malaria incidence in Magude',
+       subtitle = 'Observed (red) vs. predicted (black)',
+       x = 'Date',
+       y = 'Cases per 1,000') +
+  theme_maltem()
+```
+
+# Effect of insecticide-treated bednets (ITN)
+
+At first glance, ITN coverage appears associated with lower malaria incidence
+
+```{r}
+ggplot(data = df %>% filter(disease == 'MALARIA'),
+       aes(x = itn_coverage,
+           y = pk)) +
+  geom_point(alpha = 0.3) +
+  theme_maltem() +
+  labs(x = 'ITN coverage',
+       y = 'Weekly incidence per 1,000',
+       title = 'ITN and malaria',
+       subtitle = 'One point per district, week, age-group') +
+  scale_x_log10()
+```
+
+# Effect of indoor residual spraying (IRS)
+
+```{r}
+x <- df %>%
+  filter(disease == 'MALARIA') %>%
+  group_by(date, district, irs_coverage_people, weeks_since_last_irs_campaign_end) %>%
+  summarise(p = sum(cases) / sum(population) * 1000) %>%
+  ungroup %>%
+  mutate(irs_time_since = ifelse(weeks_since_last_irs_campaign_end <= 10,
+                                 '00-10',
+                                 ifelse(weeks_since_last_irs_campaign_end <= 50,
+                                        '11-50',
+                                        ifelse(weeks_since_last_irs_campaign_end <= 99,
+                                               '51-99',
+                                               ifelse(weeks_since_last_irs_campaign_end > 100,
+                                                      '99+',
+                                                      NA))))) %>%
+  filter(!is.na(weeks_since_last_irs_campaign_end)) %>%
+  mutate(irs_coverage = ifelse(irs_coverage_people < 50,
+                               '00-50%',
+                               ifelse(irs_coverage_people <= 70,
+                                      '51-70%',
+                                      ifelse(irs_coverage_people <= 90,
+                                             '71-90%',
+                                             ifelse(irs_coverage_people > 90,
+                                                    '91+%',
+                                                    '00%')))))
+
+ggplot(data = x,
+       aes(x = weeks_since_last_irs_campaign_end,
+           y = p)) +
+  geom_point(alpha = 0.3) +
+  facet_wrap(~irs_coverage) +
+  labs(x = 'Weeks since last IRS campaign end',
+       y = 'Incidence per 1,000',
+       title = 'Effect of IRS',
+       subtitle = 'Faceted by coverage percentage') +
+  theme_maltem()
+
+```
+
+# Incorporating more variables
+
+Here's what happens when we model cases using ITN, IRS, time since IRS campaign, seasonality, etc.
+
+```{r}
+model_data <- 
+  df %>%
+  filter(disease == 'MALARIA') %>%
+  left_join(coords@data) %>%
+  mutate(season = ifelse(month %in% 1:4, 'rainy', 'not-rainy')) %>%
+  mutate(month = factor(month)) %>%
+  group_by(date, year, 
+           season,
+           month,
+           irs_coverage_people,
+           weeks_since_last_irs_campaign_end,
+           itn_coverage,
+           district, distance_to_south_africa) %>%
+  summarise(p = sum(cases) / sum(population) * 1000) %>%
+  ungroup
+
+fit <- lm(p ~ year*log(distance_to_south_africa)  + year + season*year  + month +  irs_coverage_people*weeks_since_last_irs_campaign_end + itn_coverage + district, 
+          data = model_data %>%
+            filter(date <= '2016-06-01',
+                   date >= '2012-06-01'))
+
+x <- model_data %>% filter(district == 'MAGUDE')
+x$predicted <- predict(fit, x)
+
+ggplot() +
+  geom_line(data = x,
+       aes(x = date,
+           y = predicted),
+       alpha = 0.8) +
+  geom_line(data = x,
+       aes(x = date,
+           y = p),
+       color = 'red',
+       alpha = 0.8) +
+  labs(title = 'Weekly malaria incidence in Magude',
+       subtitle = 'Observed (red) vs. predicted (black)',
+       x = 'Date',
+       y = 'Cases per 1,000') +
+  theme_maltem()
+
+```
+
 # Questions
@@
 1. What are the mathematics of the method by which we assign weights to synthetic control districts? Are we concerned that so few districts get any weight?
+
+2. Given that proximity to South Africa seems to have an interactive effect with the decline in malaria incidence, should we include this in our model? If so, how?
+
+3. Are we currently excluding Manhiça from our synthetic controls? If so, why?
+
+4. What are we doing with IRS / ITN?
+
+5. How are we treating secular trends (ie, near linear decline in springtime malaria in Magude for the 4 years from 2012-2015)?

### 2017-07-19 get_irs_data.R

@@ -195,3 +195,27 @@ if('irs_data_cleaned.RData' %in% dir('data/binaries')){
   save(irs,
        file = 'data/binaries/irs_data_cleaned.RData')
 }
+
+# Come up with protection function
+irs_protect <- function(weeks = 1:52){
+  # Conjure up an effectiveness curve
+  the_curve <- 1 - ((weeks)^(5) / (52^(5)))
+  the_curve <- ifelse(the_curve > 1, 1, 
+                      ifelse(the_curve < 0, 0, the_curve))
+  the_curve <- the_curve
+  return(the_curve)
+}
+
+
+# 
+# x <- data.frame(weeks = 1:52,
+#                 protection = irs_protect() * 100)
+# ggplot(data = x,
+#        aes(x = weeks,
+#            y = protection)) +
+#   geom_area(alpha = 0.6,
+#             fill = 'darkorange') +
+#   labs(x = 'Weeks',
+#        y = 'Protection (%)',
+#        title = 'IRS protection over time',
+#        subtitle = 'Y-axis: percentage of maximum coverage rate')

### 2017-07-19 master.R

@@ -67,17 +67,76 @@ irs <- irs %>%
                 irs_coverage_houses,
                 irs_coverage_people)
 
+# Get which number irs campaign
+cum_summer <- function(x){
+  x[is.na(x)] <-0
+  cumsum(x)
+}
+irs <- irs %>%
+  arrange(district, year, week) %>%
+  mutate(dummy = ifelse(weeks_since_last_irs_campaign_end == 1, 1, 0)) %>%
+  group_by(district) %>%
+  mutate(irs_campaign = cum_summer(dummy))  %>%
+  mutate(irs_campaign = ifelse(weeks_since_last_irs_campaign_end == 0, 
+                               irs_campaign + 1,
+                               irs_campaign))
+
+# Create a protection variable based on decline
+irs <- irs %>%
+  mutate(irs_protection = irs_protect(weeks_since_last_irs_campaign_end) * irs_coverage_people) %>%
+  mutate(irs_protection = ifelse(irs_protection < 0, 0,
+                                 ifelse(irs_protection > 100, 100,
+                                        irs_protection))) %>%
+  mutate(irs_protection = ifelse(is.na(irs_protection), 0, irs_protection))
+
+# Remove unecessary variables
+irs <- irs %>%
+  dplyr::select(-dummy, -irs_campaign)
+
+# # During an IRS campaign increase protection linearly (not working yet)
+# x <-
+#   irs %>%
+#   # filter(!is.na(irs_campaign)) %>%
+#   arrange(year, week) %>%
+#   group_by(district, irs_campaign) %>%
+#   mutate(irs_protection = ifelse(is.na(irs_protection),
+#                                  dplyr::first(irs_protection[!is.na(irs_protection)]),
+#                                  irs_protection)) %>%
+#   mutate(irs_protection_campaign_start = dplyr::first(irs_protection[weeks_since_last_irs_campaign_end == 0]),
+#          irs_portection_campaign_end = dplyr::first(irs_protection[weeks_since_last_irs_campaign_end == 1])) %>%
+#   mutate(irs_protection_campaign_weeks = length(which(weeks_since_last_irs_campaign_end == 0))) %>%
+#   mutate(irs_protection_increase = irs_portection_campaign_end - irs_protection_campaign_start) %>%
+#   mutate(irs_protection_increase = ifelse(irs_protection_increase < 0,
[…]

### 2017-07-19 reports/misc_analysis/get_distance_to_border.R

@@ -1,0 +1,55 @@
+# Interpolate for our locations
+library(sp)
+x <- cism::moz2
+x <- x[x@data$NAME_1 %in% c('Maputo', 'Gaza'),]
+coords <- data.frame(coordinates(x))
+names(coords) <- c('x', 'y')
+coords$district <- x@data$NAME_2
+
+# Standardize names
+coords$district <- toupper(coords$district)
+coords$district[coords$district == 'CHÓKWÈ'] <- 'CHOKWE'
+coords$district[coords$district == 'GUIJÁ'] <- 'GUIJA'
+coords$district[coords$district == 'XAI-XAI'] <- 'XAI-XAI DISTRICT'
+coords$district[coords$district == 'MANHIÇA'] <- 'MANHICA'
+coords$district[coords$district == 'MATUTUÍNE'] <- 'MATUTUINE'
+coords$district[coords$district == 'MANDLAKAZI'] <- 'MANJACAZE'
+
+# Add a row for matola
+matola <- data_frame(x = 32.46580,
+                     y = -25.92391,
+                     district = 'MATOLA')
+coords <- bind_rows(coords, 
+                    matola)
+# Add a row for xai-xai city
+xxc <- coords %>% filter(district == 'XAI-XAI DISTRICT') %>%
+  mutate(district = 'XAI-XAI CITY')
+coords <- bind_rows(coords,
+                    xxc)
+# Make spatial
+coords <- coords %>%
+  mutate(lng = x,
+         lat = y)
+coordinates(coords) <- ~x+y
+proj4string(coords) <- proj4string(cism::moz3)
+
+# Get South Africa and Swaziland
+sa <- cism::africa
+sa <- sa[sa@data$COUNTRY %in% c('South Africa', 'Swaziland'),]
+
+# Get projected versions
+sa_proj <- spTransform(sa,
+                       CRS("+init=epsg:3347"))
+coords_proj <- spTransform(coords,
+                           CRS("+init=epsg:3347"))
+
+# Get distance to south Africa
+coords@data$distance_to_land_border <- NA
+for (i in 1:nrow(coords)){
+  this_distance <- rgeos::gDistance(coords_proj[i,], sa_proj)
+  coords@data$distance_to_land_border[i] <- this_distance / 10000
+}
+distances_to_border <- coords@data
+distances_to_border <-
+  distances_to_border %>%
+  dplyr::select(-lat, -lng)

### 2017-08-18 paper.Rmd
london docs

--- 2017-03-09 paper.Rmd
@@
 title: "Cost-effectiveness of a malaria elimination campaign"
-author: "Laia Cirera, Joe Brew, Elisa Sicuri"
-date: "`r Sys.Date()`"
+subtitle: "Cross-software reproducibility check"
+author: "Joe Brew, Laia Cirera, Ranjeeta Thomas, Elisa Sicuri"
 output:
@@
 knitr::opts_chunk$set(comment = NA, 
-               echo = FALSE, 
+               echo = TRUE, 
                warning = FALSE, 
@@
 ```{r}
-library(readxl)
-library(tidyverse)
-library(broom)
-library(ggthemes)
-
-# Read in costs database 
-costs <- read_excel("public_data/costs.xlsx")
-
-# Read in epi parameters 
-parameters <- read_excel("public_data/parameters.xlsx")
-
-# Read in incidence data and weatehr data
-source("master.R")
-bes <- df
-
-# Create variables 
-bes <- 
-  bes %>%
-  mutate(mda=ifelse(date<"2015-11-08", "pre_mda", 
-                    ifelse(date<"2016-02-04", "mda", "post_mda")),
-         intervention=district=="MAGUDE",
-         magude = ifelse(district == 'MAGUDE', 'Magude', 'Not Magude'))
-
-# Create model to estimate effectiveness
-mod <-lm(pk ~ mda*intervention + irs_coverage, data=bes, weights = population)
+source('master.R')
 ```
 
-Model parameters
+```{r, message = FALSE, results = 'hide'}
+# Get packages
+library(Synth)
+
+# Prepare data
+gen_dummy <- function(condition){ifelse(condition, 1, ifelse(!condition, 0, NA))}
+
+model_data <- 
+  df %>%
+  filter(age_group == '0-4') %>%
+  filter(disease == 'MALARIA',
+         district != 'XAI-XAI CITY',
+         district != 'MANHICA',
+         district != 'MASSINGIR')
+
+# Get a standardized time variable
+model_data <-
+  model_data %>%
+  mutate(time = as.Date(paste0(year, '-01-01')) + (week*7)) %>%
+  mutate(time = as.numeric(time)) %>%
+  mutate(high_season = ifelse(month %in% 1:3, 1, 0))
+
+# Generate a oct-dec lag
+high_season <- model_data %>%
+  filter(high_season == 1) %>%
+  group_by(district, year) %>%
+  summarise(cases = mean(pk)) %>%
+  ungroup %>%
+  group_by(district) %>%
+  mutate(cases_lag = dplyr::lag(cases, n = 1, default = NA)) %>%
+  ungroup %>%
+  dplyr::select(-cases)
+model_data <-
+  left_join(model_data,
+            high_season)
+# Get 1 week lags too
+model_data <-
+  model_data %>%
+  arrange(date) %>%
+  group_by(district) %>%
+  mutate(cases_lag1 = dplyr::lag(pk, n = 1, default = NA),
+         cases_lag2 = dplyr::lag(pk, n = 2, default = NA))
+
+# Filter away some stuff
+model_data <- 
+  model_data %>%
+  filter(week < 52) %>%
+  filter(date >= '2013-10-01')
+
+# Get data with dummies
+model_formula <- lm(pk ~  high_season + year + time, data = model_data)
+model_matrix <- data.frame(model.matrix(model_formula)) %>%
+  dplyr::select(-`X.Intercept.`) %>%
+  mutate(pk = model_data$pk,
+         district = model_data$district,
+         itn_coverage = model_data$itn_coverage) %>%
+  mutate(district_name = district) %>%
+  mutate(district = as.numeric(factor(district)),
+         cases_lag = model_data$cases_lag,
+         cases_lag1 = model_data$cases_lag1,
+         cases_lag2 = model_data$cases_lag2)
+
+left <- 
+  expand.grid(district = sort(unique(model_matrix$district)),
+              time = sort(unique(model_matrix$time)))
+right <- 
+  model_matrix %>%
+  group_by(district, time) %>%
+  tally
+balanced <- left_join(left, right)
+model_matrix <-
+  left_join(left, model_matrix)
+
+# Fill the NAs
+model_matrix <-
+  model_matrix %>%
+  arrange(time) %>%
+  group_by(district) %>%
+  mutate(high_season = zoo::na.locf(high_season), 
+         year = zoo::na.locf(year),
+         pk = zoo::na.locf(pk),
+         itn_coverage = zoo::na.locf(itn_coverage),
+         district_name = zoo::na.locf(district_name))
+
+model_matrix <- data.frame(model_matrix)
+
+prior_time <- sort(unique(model_matrix$time))
+prior_time <- prior_time[prior_time <= as.numeric(as.Date('2015-08-01'))]
+prepared_data <- 
+  Synth::dataprep(foo = model_matrix,
+                  predictors = c('itn_coverage',
+                                 'cases_lag',
+                                 'cases_lag1',
+                                 'cases_lag2'),
+                  predictors.op = 'mean',
+                  dependent = 'pk',
+                  time.variable = 'time',
+                  unit.variable = 'district',
+                  treatment.identifier = which(sort(unique(model_data$district)) == 'MAGUDE'),
+                  controls.identifier = which(sort(unique(model_data$district)) != 'MAGUDE'),
+                  unit.names.variable = 'district_name',
+                  time.predictors.prior = prior_time,
+                  time.optimize.ssr = prior_time,
+                  time.plot = sort(unique(model_matrix$time)))
+
+out <- Synth::synth(data.prep.obj = prepared_data)
+# synth_tab <- synth.tab(synth.res = out,
+#           dataprep.res = prepared_data,
+#           round.digit = 3)
+# # Predictor weights
+# out$solution.v
+# # Unit weights
+# out$solution.w
+```
+
+# Unit weights
 
 ```{r}
-
-# Explore results of linear model
-# knitr::kable(tidy(summary(mod)))
+uw <- left_join(
+  model_matrix %>%
+    filter(!duplicated(district)) %>%
+    dplyr::select(district, district_name),
+  data.frame(out$solution.w) %>%
+    mutate(district = as.numeric(row.names(.)))
+)
+uw
+ggplot(data = uw,
+       aes(x = district_name,
+           y = w.weight)) +
+  geom_bar(stat = 'identity',
+           alpha = 0.6,
+           fill = 'darkorange') +
+  theme(axis.text.x = element_text(angle = 90)) +
+  labs(x = 'District',
+       y = 'Weight',
+       title = 'Weights by district') +
+  ggthemes::theme_hc()
 ```
 
+# Comparison plot
+
+(Difference between synthetic and observed)
+
 ```{r}
-# Use model to predict conterfactual
-other_irs <- bes %>% filter(year == 2016, district != "MAGUDE") %>% summarise(x =mean(irs_coverage, na.rm =TRUE)) %>% .$x
-counter_factual <- 
-   bes %>% filter(year == 2016, 
-                  district == "MAGUDE")
-results <- list()
-for (i in 2010:2020){
-  new_year <- counter_factual %>%
-    mutate(year = i) 
-  results[[i]] <- new_year
-}
-counter_factual <- bind_rows(results)
-counter_factual$date <- 
-  as.Date(paste0(counter_factual$year, substr(counter_factual$date, 5, 10)))
-
-counter_factual <- 
-  counter_factual %>%
-  mutate(irs_coverage = other_irs) %>%
-  mutate(intervention = FALSE)
-
-counter_factual$p_cf <- predict(mod, counter_factual)
-counter_factual$cases_cf <- counter_factual$p_cf*counter_factual$population 
-counter_factual$cases_averted <- counter_factual$cases_cf - counter_factual$cases
-
-# Get sum of averted cases
-ac <- counter_factual %>%
-  filter(year == 2016) %>%
-  group_by(age)%>%
-  summarise(sum_cases_averted=sum(cases_averted))
-
-# Get the reduction in incidence
-inc <-
-  counter_factual %>%
-  filter(year == 2016) %>%
-  group_by(age)%>%
-  summarise(reduction = (mean(p_cf) - mean(p)) / mean(p_cf) * 100)
-
-### Create new variables in our epi data related to clinical events
-counter_factual <- 
-  counter_factual %>%
-  mutate(cases_non_severe = cases * parameters$value[parameters$parameter == 'p_non_sev'],
-         cases_non_severe_cf = cases_cf * parameters$value[parameters$parameter == 'p_non_sev'],
-         cases_severe = cases * parameters$value[parameters$parameter == 'p_sev'],
-         cases_severe_cf = cases_cf * parameters$value[parameters$parameter == 'p_sev'],
-         deaths = cases * parameters$value[parameters$parameter == 'cfr'],
-         deaths_cf = cases_cf * parameters$value[parameters$parameter == 'cfr'],
-         cost_outpatient = cases_non_severe * parameters$value[parameters$parameter == 'cost_outp_usd'],
-         cost_outpatient_cf = cases_non_severe_cf * parameters$value[parameters$parameter == 'cost_outp_usd'],
-        cost_inpatient = cases_severe * parameters$value[parameters$parameter == 'cost_inp_usd'],
-        cost_inpatient_cf = cases_severe_cf * parameters$value[parameters$parameter == 'cost_inp_usd']) %>%
-  mutate(cost_clinical = cost_inpatient + cost_outpatient,
-         cost_clinical_cf = cost_inpatient_cf + cost_outpatient_cf)
-
-# Dollars per case averted
-# We have cases_averted
-# We also have clinical costs saved
-# We need program costs
-
-# 2016
-# dpca = program costs + real clinical costs / cf program costs + cf clinical costs
-
-# We need to differentiate between program costs (MALTEM) and intervention costs
-intervention_costs <- costs[!costs$elimination_activity %in% c('Entomology', 'REACT'),]
-
-intervention_costs <-
-  intervention_costs %>%
-  # MALTEM COSTS
-  mutate(cost_maltem = 
-              (resource_use * 
-              q_maltem *
-              p_maltem *
-              yearly_regularity / 
-              useful_life)) %>%
-  mutate(transport_meticais = ifelse(transport_cost, 0.25 * cost_maltem, 0),
-         import_meticais = ifelse(import_tax, 0.4 * cost_maltem, 0),
-         operating_meticais = ifelse(operating_cost, 0.1 * cost_maltem, 0),
-         maintenance_meticais = ifelse(maintenance_cost, 0.15 * cost_maltem, 0)) %>%
-  mutate(cost_maltem = 
-           cost_maltem +
-           transport_meticais +
-           import_meticais +
-           operating_meticais +
-           maintenance_meticais) %>%
-  mutate(cost_basic = 
-           (resource_use * 
-              q_basic *
-              (p_basic_perc * p_maltem) *
-              yearly_regularity / 
-              useful_life)) %>%
-    mutate(transport_meticais_basic = ifelse(transport_cost, 0.25 * cost_maltem, 0),
-         import_meticais_basic = ifelse(import_tax, 0.2 * cost_maltem, 0),
-         operating_meticais_basic = ifelse(operating_cost, 0.1 * cost_maltem, 0),
-         maintenance_meticais_basic = ifelse(maintenance_cost, 0.15 * cost_maltem, 0)) %>%
-    mutate(cost_basic = 
-           cost_basic +
-           transport_meticais_basic +
-           import_meticais_basic +
-           operating_meticais_basic +
-           maintenance_meticais_basic) %>%
-  dplyr::select(elimination_activity,
-                budget_category,
-                baseline_year,
-                item,
-                description,
-                cost_basic,
-                cost_maltem) %>%
-  mutate(cost_basic = ifelse(is.na(cost_basic), 0, cost_basic))
-
-# Get total intervention costs
-total_intervention_cost <- 
-  intervention_costs[,c('cost_basic', 'cost_maltem')]
-total_intervention_cost <-
-  tidy(apply(total_intervention_cost, 2, sum, na.rm = TRUE) / 75)
-
-# Get a counterfactula for government
-cf_costs <- intervention_costs %>%
-  filter(elimination_activity %in% c('IRS', 'LLIN')) %>%
-  summarise(x = sum(cost_basic) / 75)
-cf_costs <- data_frame(names = 'counter_factual',
-                       x = cf_costs$x)
-
-total_intervention_cost <-
-  bind_rows(total_intervention_cost,
-            cf_costs)
-names(total_intervention_cost)[2] <- 'cost'
-
-ggplot(data = total_intervention_cost,
-       aes(x= names,
-           y = cost)) +
-  geom_bar(stat = 'identity') +
-  labs(x = 'Scenario',
-       y = 'US Dollars',
-       title = 'Comparative costs of malaria control/elimination approaches',
-       subtitle = 'USD') +
-  geom_label(aes(label = round(cost)))
-
-# Add clinical costs to our total_intervention_cost dataset
-clinical_costs <- 
-  counter_factual %>%
-  filter(year == 2016) %>%
-  summarise(counter_factual = sum(counter_factual$cost_clinical_cf),
-            cost_basic = sum(counter_factual$cost_clinical),
-            cost_maltem = sum(counter_factual$cost_clinical))
-clinical_costs <- tidy(t(clinical_costs))
-names(clinical_costs) <- c('names', 'clinical_cost')
-total_intervention_cost <-
-  left_join(x = total_intervention_cost,
-            y = clinical_costs,
-            by = 'names')
-
-# Get total cost (program plus clinic)
-total_intervention_cost$total_cost <-
-  total_intervention_cost$cost +
-  total_intervention_cost$clinical_cost
-
-# Get total cases for 2016 under both scenarios
-x <- counter_factual %>%
-  filter(year == 2016) %>%
-  summarise(cases_maltem = sum(cases),
-            cases_cf = sum(cases_cf))
-
-DT::datatable(total_intervention_cost)
+path.plot(synth.res = out,
+          dataprep.res = prepared_data)
+abline(v = as.numeric(as.Date('2015-10-01')), col = 'red')
 ```
 
-# Introduction
+# Gaps plot
 
-Mozambique is one of the 10 highest-malaria burden countries in the world, with parasite prevalence ranging from 3 to more than 50%.
-Although large drops in transmission can be achieved in a limited timeframe, indoor residual spraying and case management alone are not sufficient to attain elimination (Okumu et al, 2011).
-
-Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investiga??o em Sa?de de Manhi?a (CISM).
-The main objective of this program is to assist the NMCP in adopting targeted evidence-based elimination plans through a 'learn by doing' strategy that begins in program mode and incorporates data from rigorous evaluation-into action as it is continuously being generated.
-Besides developing a fully costed, targeted, evidence-based elimination plan for Southern Mozambique, the first two years of this program have generated knowledge to inform the programmatic aspects of the elimination plan, most importantly how to interrupt transmission.
-It has also provided coordination, management and capacity building support to the NMCP, and facilitated the adoption of evidence-based malaria policies through the creation of a Malaria Technical Advisory Committee in the country.
-
-The  MALTEM program will provide technical, operational and financial support to design and scale up new strategies for the Mozambican National Malaria Control Program (NMCP) to achieve elimination in Maputo and significantly decrease malaria transmission rates in Gaza and Inhambane provinces by 2020.
-
-Within this context, a rigorous costing analysis needs to be carried out in order to assess the viability of MALTEM strategies and guarantee the financial sustainability of the project across time.
-Firstly, this study estimates the costs associated with malaria elimination activities in the district of Magude and models these costs, in accordance to economic theory principles, evidence from literature on malaria elimination as well as from WHO guidelines, across time.
-These costs are then compared with malaria control costs to identify potential savings of resources associated with malaria elimination, in the medium/long run.
-The whole analysis allows us assessing whether it malaria elimination (by comparing the costs of malaria elimination vs.
-costs of malaria control activities implementation and estimating the quantity and value of additional resource requirements to achieve malaria elimination), is financially affordable and sustainable over the next 10 years, but also, to which extend malaria elimination is also cost-effective compared to malaria control interventions.
-
-#Methods
-
-##Study area and population
-
-The malaria elimination is taking place in Magude District, on the North-Western area of Maputo Province, Southern Mozambique.
-Magude District borders with the district of Massingir (Gaza Province), in the North, Manhi?a District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West.
-It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds.
-The district has seven health facilities with a total of 43 beds ^[(1.65 beds per 1,000 inhabitants)]  and reported 20,000 malaria cases per 100,000 population at risk in 2012.
-
-Manhi?a district has been used as a comparator area for the analysis of the interventions, costing and variables of interest.
-Manhi?a, located 80km north of the capital, Maputo, is a semi-rural area, with a population of 165.497 habitants.
-As of 2016, malaria remains one of the main causes of under-five morbidity and mortality in the district.
-
-##Study design
-
-The study we propose is not a randomized one.
-Malaria elimination activities will be carried out in Magude and not in a randomized way (it is not possible to randomize a district based malaria elimination campaign).
-In order to identify the impact of malaria elimination activities on indicators of impact and in order to perform the general cost-effectiveness analyses, a comparator is needed.
-There is rich data existent in Manhi?a district, where CISM is located.
-However, it is subject to important noise given all biomedical research and clinical trials carried out within the district in the last 20 years.
-In consequence, we will use and ponderate data at the district level.
-
-In the estimates we will also control for potential confounders, because good data are available.
-It is possible to control, for instance, for the demographic structure of people living in the districts, daily rainfall and temperature data, etc.
-Features of the health systems of all districts in the province are also known.
-
-##Costs
-
-###Costs of malaria elimination
-
-The costing of the malaria elimination plan in Magude District consists of a detailed compilation of the estimated costs of resources used in the first 2 years and the additional resources needed to carry out the 5 year Malaria elimination program (till 2020).
-The main purpose of this costing analysis is to provide realistic information about the financial and economic resources used to execute the plan as well as the distribution of these resources across the years......
-
-###Costs of malaria case management
-
-ELISA's ref.
-Health system costs for the treatment of a malaria episode included admission costs at the maternity ward in the case of inpatients, or the cost of attendance and treatment in the case of outpatients.
-Unit costs of an admission at the maternity ward (occupied bed/day) and of an external consultation included all recurrent components and excluded capital costs.
-Recurrent costs consisted of personnel, medical, surgical and laboratory supplies among other recurrent costs.
-Unit costs were updated from the year 2000 to the year 2007 using the average annual rate inflation correction
-factor and validated during interviews with the administrative staff of the MDH [32].
-Drug costs were added to this estimate.
-Quinine was the antimalarial drug administered to admitted pregnant women and to outpatient pregnant women in the first trimester of gestation.
-The combination of SP and artesunate was administered to outpatient pregnant women who were over the first trimester of pregnancy.
-Drug costs for malaria treatment are mostly sustained by the National Health System in Mozambique, and patients only pay a small fee when attending government health facilities.
-Total admission costs were calculated by multiplying the daily unit cost of admission at the maternity ward by the average number of admission days, which were estimated through the household cost
-data collection outlined below.
-
-##Costing scenarios and sensitivity analysis
-
-Explain costs of malaria control
-
-ELISA's ref.
-The base case costing scenario relied on the following set of assumptions: a discount rate of 3 % was applied to capital costs; wastage of RDT kits, ACT treatment courses and other field work materials was assumed to be 10 %; overhead for national (and international) supervision amounted to 15 % of the total direct financial costs of the programme.
-These are believed to be conservative assumptions.
-The cost of ACT treatment courses and RDT kits
-were based on the cost, insurance and freight (cif.) price of the drug or diagnostic derived ......
-
-Elisa's ref.
-Probabilistic cost-effectiveness analysis was undertaken through Monte Carlo simulations using @Risk (version 5.0) add-in tool to Microsoft Excel.
-(Palisade Corporation, Ithaca, NY, USA).
-The main factors affecting the cost-effectiveness of the intervention were identified in the estimated model by calculating Spearman's rank correlation coefficients as a measure of the magnitude of the association between each variable and the ICERs.
-A threshold analysis of the cost-effectiveness of IPTp-SP was performed to estimate cut-off points beyond which the
-prevention is no longer cost-effective.
-The threshold level of the ICERs used to define the intervention as cost-effective was 129 US$ per DALY averted, while 36 US$ per DALY averted was the threshold used to define the intervention as highly cost-effective.
-These threshold figures were based on previously accepted 1993 and 1996 World Bank definitions and inflated to their 2007 equivalent: 129 and 36 US$ for the year 2007 correspond to 100
-and 25 US$ for the year 1993, respectively [47-49].
-The costeffective intervention threshold was conservatively set at 100 instead of 150 US$ per DALY averted.
-Probabilistic threshold analysis was performed on SP price, other intervention costs, case fatality rate, malaria incidence, protective efficacy and antenatal clinic attendance.
-Furthermore, a one-way sensitivity analysis was undertaken on SP price.
-
-##Effectiveness
-
-All explanation on data from BES, its limitations and advantages compared to data from MALTEM's own surveillance system.......
+(Same as above, but showing the gap between synthetic and real, rather than the actual values)
 
 ```{r}
-x <- counter_factual
-x$predicted <- predict(mod, x)
-long <- gather(x, key, value, cases:predicted)
-long$value <- as.numeric(long$value)
-long <- long %>%
-  mutate(value = ifelse(key == 'pk', NA, value))
-
-y <- bes
-long2 <- gather(y, key, value, cases:predicted)
-long2$value <- as.numeric(long2$value)
-
-z <- rbind(long %>%
-             filter(year >= 2017),
-           long2 %>%
-             filter(year <= 2016))
-
-ggplot(data = z %>%
-         filter(key %in% c('predicted', 'pk'),
-                district == 'MAGUDE'),
-       aes(x = date,
-           y = value,
-           group = key,
-           color = key)) +
-  geom_line() +
-  facet_wrap(~age)
+gaps.plot(synth.res = out,
+          dataprep.res = prepared_data)
+abline(v = as.numeric(as.Date('2015-10-01')), col = 'red')
 ```
-
-```{r}
-# Plot the observed and counterfactual
-plot_data <- counter_factual %>%
-  group_by(date) %>%
-  summarise(cases = sum(cases),
-            cases_cf = sum(cases_cf),
-            cases_averted = sum(cases_averted)) %>%
-  gather(key, value, cases:cases_averted) %>%
-  filter(key != "cases_averted") %>%
-  mutate(key = ifelse(key == "cases", "MALTEM", "Counter-factual"))
-
-ggplot(data=plot_data,
-       aes(x=date,
-           y=value,
-           color=key))+
-  geom_line(size =2,
-            alpha = 0.6) +
-  labs(title="Observed malaria cases vs counterfactual",
-       subtitle="Post-MDA period, 2016",
-       x="Date",
-       y="Malaria cases") +
-  theme_minimal() +
-  scale_color_manual(name = "",
-values = c("red", "darkgreen"))
-```
-
-##Cost effectiveness analysis
-
-####Cost-effectiveness of MALTEM on malaria
-
-###Cost-effectiveness of MALTEM on mortality
-
-#Results
-
-##Costs
-
-###Costs elimination (baseline years)
-
-###Costs projection elimination
-
-###Costs projection control
-
-###Costs savings elimination
-
-###Effectiveness (impact)
-
-###Outcome indicators (costing outcomes)
-
-###Impact indicators
-
-###Cost-effectiveness analysis
-
-And here is another figure.
-
-```{r, fig.height = 8}
-
-ggplot(data = bes,
-       aes(x = date,
-           y = pk)) +
-  geom_line(aes(color = age)) +
-  facet_wrap(~district) +
-  labs(x = 'Date',
-       y = 'Incidence (cases per 1,000)',
-       title = 'Malaria incidence over time by district',
-       subtitle = 'Cases per 1,000 inhabitants') +
-  scale_color_manual(name = 'Age',
-                     values = c('darkgreen', 'darkorange'))
-```
-
-```{r}
-x <- bes %>%
-    filter(district != 'NAMAACHA') %>%
-  # filter(district != 'MANHICA') %>%
-  group_by(date,
-           # year, month,
-           # age, 
-           magude) %>%
-  summarise(#pk = weighted.mean(pk, w = population),
-            cases = sum(cases),
-            pop = sum(population),
-            precip = mean(precip)) %>%
-  ungroup %>%
-  mutate(p = cases / pop * 1000) #%>%
-  # mutate(date = as.Date(paste0(year, '-',
-  #                              '06',
-  #                              # month, 
-  #                              '-01')))# %>%
-  # filter(date >= '2014-07-01')
-
-library(ggthemes)
-library(cism)
-ggplot(data = x,
-       aes(x = date,
-           color = magude,
-           y = p)) +
-  geom_line(alpha = 0.6) +
-  # geom_smooth() +
-  # facet_wrap(~age, ncol = 1) +
-  # theme_bw()  +
-  scale_color_manual(name = '',
-                     values = c('blue', 'red')) +
-  # theme_fivethirtyeight() +
-  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-12-31'))),
-             lty = 2, alpha = 0.2) +
-  geom_vline(xintercept = as.numeric(as.Date(paste0('2015-03-15'))),
-             color = 'darkgreen') +
-  labs(x = 'Week',
-       y = 'Weekly incidence per 1,000',
-       title = 'Weekly clinical malaria',
-       subtitle = 'Province of Maputo, all ages, weekly') +
-  geom_label(data = data_frame(x = as.Date(paste0('2015-03-15')),
-                               y = 15,
-                               label = 'MDA'),
-             aes(x = x,
-                 y = y,
-                 label = label),
-             color = 'black') +
-  theme_cism()
-```
-# Discussion
-
-```{r}
-bes <- bes %>% filter(district != 'NAMAACHA')
-# Prediction
-bes$predicted <- predict(mod, bes)
-
-x <- bes %>%
-    filter(district != 'NAMAACHA') %>%
-  # filter(district != 'MANHICA') %>%
-  group_by(date,
-           # year, month,
-           # age, 
-           magude) %>%
-  summarise(#pk = weighted.mean(pk, w = population),
-            cases = sum(cases),
-            pop = sum(population),
-            precip = mean(precip),
-            precip_risk = mean(precipitation_risk),
-            predicted = mean(predicted)) %>%
-  ungroup %>%
-  mutate(p = cases / pop * 1000)
-#%>%
-  # mutate(date = as.Date(paste0(year, '-',
-  #                              '06',
-  #                              # month, 
-  #                              '-01')))# %>%
-  # filter(date >= '2014-07-01')
-
-library(ggthemes)
-ggplot(data = x,
-       aes(x = date,
-           color = magude,
-           y = predicted)) +
-  geom_line(alpha = 0.6) +
-  # geom_smooth() +
-  # facet_wrap(~age, ncol = 1) +
-  # theme_bw()  +
-  scale_color_manual(name = '',
-                     values = c('blue', 'red')) +
-  theme_fivethirtyeight() +
-  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-12-31'))),
-             lty = 2, alpha = 0.6) +
-  geom_vline(xintercept = as.numeric(as.Date(paste0('2015-03-15')))) +
-  geom_line(data = x,
-            aes(x = date,
-                y = p,
-                color = magude),
-                lty = 2,
-            alpha = 0.5)
-
-```
-
-```{r}
-# Precipitation
-x <- bes %>%
-  group_by(year, month) %>%
-  summarise(p = mean(precipitation)) %>%
-  mutate(date = as.Date(paste0(year, '-', month, '-01')))
-ggplot(data = x,
-       aes(x = date,
-           y = p)) +
-  geom_bar(stat = 'identity') 
-```

### 2017-08-18 master.R

@@ -165,21 +165,21 @@ df_agg <- df %>%
             wind_speed_max = first(wind_speed_max),
             distance_to_land_border = first(distance_to_land_border))
 
-# # Write data for sharing with collaborators
-write_csv(df_agg, 'data/outputs/cases_population_weather_itn_irs_pooled_age_groups.csv')
-write_csv(df_young, 'data/outputs/cases_population_weather_itn_irs_young_only.csv')
-write_csv(df_old, 'data/outputs/cases_population_weather_itn_irs_old_only.csv')
-write_csv(df, 'data/outputs/cases_population_weather_itn_irs.csv')
-write_csv(bes, 'data/outputs/cases_only.csv')
-write_csv(pop, 'data/outputs/population.csv')
-# Write csv for weather data
-write_csv(weather_weekly, 'data/outputs/weather_weekly.csv')
-write_csv(weather, 'data/outputs/weather_daily.csv')
-# write_csv(wide_weather, 'data/outputs/weather_wide.csv')
-# Write csv for itn data
-write_csv(itn, 'data/outputs/itn.csv')
-# Write csv for irs data
-write_csv(irs, 'data/outputs/irs.csv')
+# # # Write data for sharing with collaborators
+# write_csv(df_agg, 'data/outputs/cases_population_weather_itn_irs_pooled_age_groups.csv')
+# write_csv(df_young, 'data/outputs/cases_population_weather_itn_irs_young_only.csv')
+# write_csv(df_old, 'data/outputs/cases_population_weather_itn_irs_old_only.csv')
+# write_csv(df, 'data/outputs/cases_population_weather_itn_irs.csv')
+# write_csv(bes, 'data/outputs/cases_only.csv')
+# write_csv(pop, 'data/outputs/population.csv')
+# # Write csv for weather data
+# write_csv(weather_weekly, 'data/outputs/weather_weekly.csv')
+# write_csv(weather, 'data/outputs/weather_daily.csv')
+# # write_csv(wide_weather, 'data/outputs/weather_wide.csv')
+# # Write csv for itn data
+# write_csv(itn, 'data/outputs/itn.csv')
+# # Write csv for irs data
+# write_csv(irs, 'data/outputs/irs.csv')
 
 # # Join with wide weather
 # df_wide <- 

### 2017-08-18 visualizations.Rmd
running figures locally

---
title: "Publication-quality visualizations for MALTEM effectiveness paper"
subtitle: "Reproducible code and output"
author: "Brew, Cirera, Sicuri, Thomas"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

What follows is the code used for the generation of publication quality figures. Note that certain dependencies' source code are not included in this document (for example, references to other `R` scripts); in these cases, all code is available at https://github.com/joebrew/maltem_cost_effectiveness.

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               # fig.height = 5,
               # fig.width = 8,
               fig.align = 'center')

```

```{r}
# Data preparation
source('master.R')

# Define theme for plotting
# Borrowed from https://rpubs.com/Koundy/71792
theme_publication <- function(base_size=14) {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            panel.border = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(), 
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.5, "cm"),
            legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold")
    ))
}

colors <- colorRampPalette(c('orange', 'red', 'darkgrey'))(3)

scale_fill_publication <- function(...){
  library(scales)
  discrete_scale("fill",
                 "Publication",
                 manual_pal(values = colors
                              # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)

}

scale_colour_publication <- function(...){
  library(scales)
  discrete_scale("colour",
                 "Publication",
                 manual_pal(values = colors
                            # c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")
                            ), ...)

}

# Remove certain areas as per Ranjeeta's filtering
df <- 
  df %>%
  filter(disease == 'MALARIA',
         district != 'XAI-XAI CITY',
         district != 'MANHICA',
         district != 'MASSINGIR')

model_data <- df %>%
    filter(age_group == '0-4')

# Get a standardized time variable
model_data <-
  model_data %>%
  mutate(time = as.Date(paste0(year, '-01-01')) + (week*7)) %>%
  mutate(time = as.numeric(time)) %>%
  mutate(high_season = ifelse(month %in% 1:3, 1, 0))

# Generate a oct-dec lag
high_season <- model_data %>%
  filter(high_season == 1) %>%
  group_by(district, year) %>%
  summarise(cases = mean(pk)) %>%
  ungroup %>%
  group_by(district) %>%
  mutate(cases_lag = dplyr::lag(cases, n = 1, default = NA)) %>%
  ungroup %>%
  dplyr::select(-cases)
model_data <-
  left_join(model_data,
            high_season)
# Get 1 week lags too
model_data <-
  model_data %>%
  arrange(date) %>%
  group_by(district) %>%
  mutate(cases_lag1 = dplyr::lag(pk, n = 1, default = NA),
         cases_lag2 = dplyr::lag(pk, n = 2, default = NA))

# Filter away some stuff
model_data <- 
  model_data %>%
  filter(week < 52) %>%
  filter(date >= '2013-10-01')

# Get data with dummies
model_formula <- lm(pk ~  high_season + year + time, data = model_data)
model_matrix <- data.frame(model.matrix(model_formula)) %>%
  dplyr::select(-`X.Intercept.`) %>%
  mutate(pk = model_data$pk,
         district = model_data$district,
         itn_coverage = model_data$itn_coverage) %>%
  mutate(district_name = district) %>%
  mutate(district = as.numeric(factor(district)),
         cases_lag = model_data$cases_lag,
         cases_lag1 = model_data$cases_lag1,
         cases_lag2 = model_data$cases_lag2)

left <- 
  expand.grid(district = sort(unique(model_matrix$district)),
              time = sort(unique(model_matrix$time)))
right <- 
  model_matrix %>%
  group_by(district, time) %>%
  tally
balanced <- left_join(left, right)
model_matrix <-
  left_join(left, model_matrix)

```

# Map

```{r, fig.width = 8, fig.height = 5}
library(cism)
library(sp)
library(tidyverse)
library(broom)
library(maptools)
library(RColorBrewer)
library(ggrepel)

data <- moz3
data <- data[data@data$NAME_1 %in% c('Maputo', 'Gaza'),]
data@data$id <- paste0(data@data$NAME_1,
                       '|',
                       data@data$NAME_2)
data <- tidy(data, region = 'id')
data$province <- 
  unlist(lapply(strsplit(data$id, '|', fixed = TRUE), function(x){x[1]}))
data$district <- 
  unlist(lapply(strsplit(data$id, '|', fixed = TRUE), function(x){x[2]}))

label_df <- data %>%
  group_by(district) %>%
  summarise(long = mean(long),
            lat = mean(lat))

# Get centroids of provinces
centroids_p <-
  moz1@data %>%
  bind_cols(coordinates(moz1) %>% data.frame() %>% rename(x = X1, y = X2))

g1 <- ggplot() +
  theme_publication() +
  geom_polygon(data = cism::moz0_fortified,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = 'white',
               color = 'black',
               size = 0.2) +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = province),
               alpha = 0.5) +
  coord_map() +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'black',
               fill = NA,
               alpha = 0.6,
               size = 0.2) +
  scale_fill_manual(name = 'Province',
                    values = grey(c(0.05, 0.35))) +
  # geom_label_repel(data = label_df,
  #            aes(x = long,
  #                y = lat,
  #                label = district),
  #            label.padding = unit(0.05, "lines"),
  #            size = 3,
  #            alpha = 0.9) +
  geom_label_repel(data = centroids_p %>%
                     filter(NAME_1 %in% c('Gaza', 'Maputo')),
             aes(x = x,
                 y = y,
                 label = NAME_1),
             label.padding = unit(0.05, "lines"),
             box.padding = unit(3, 'lines'),
             arrow = arrow(length = unit(0.02, "npc")),
             size = 7,
             alpha = 0.9,
             segment.alpha = 0.6,
             alpha = 0.8,
             min.segment.length = unit(0, "lines")) +
  labs(x = 'Longitude',
       y = 'Latitude') +
  ggthemes::theme_map() +
  coord_map() +
  theme(legend.position = 'none',
        legend.text = element_text(size = 13)) +
  theme(plot.title = element_text(size=13),
        plot.subtitle = element_text(size = 16))

g2 <- ggplot() +
  coord_map() +
  theme_publication() +
  geom_polygon(data = data,
               aes(x = long,
                   y = lat,
                   group = group),
               color = 'black',
               fill = NA,
               alpha = 0.6,
               size = 0.2) +
  scale_fill_manual(name = 'Province',
                    values = grey(c(0.05, 0.35))) +
  # geom_label_repel(data = label_df,
  #            aes(x = long,
  #                y = lat,
  #                label = district),
  #            label.padding = unit(0.05, "lines"),
  #            size = 3,
  #            alpha = 0.9) +
  labs(x = 'Longitude',
       y = 'Latitude') +
  ggthemes::theme_map() +
  coord_map() +
  theme(legend.position = 'right',
        legend.text = element_text(size = 13)) +
  theme(plot.title = element_text(size=13),
        plot.subtitle = element_text(size = 16)) +
    geom_polygon(data = data %>%
                   filter(district == 'Magude'),
               aes(x = long,
                   y = lat,
                   group = group),
               alpha = 0.5,
               fill = grey(0.1)) +
    geom_label_repel(data = data %>%
                       filter(district == 'Magude') %>%
                       group_by(district) %>%
                       summarise(long = mean(long),
                                 lat = mean(lat)),
             aes(x = long,
                 y = lat,
                 label = district),
             label.padding = unit(0.05, "lines"),
             box.padding = unit(6, 'lines'),
             arrow = arrow(length = unit(0.02, "npc")),
             size = 7,
             alpha = 0.9,
             segment.alpha = 0.6,
             alpha = 0.8,
             min.segment.length = unit(0, "lines"),
             nudge_x = 2)

library(grid)
library(gridExtra)
g <- grid.arrange(g1, g2, top = textGrob("Geography of intervention and control areas", gp=gpar(fontsize=15, fontface = 2)), ncol = 2)

# Define functionfor siaving files
save_figs <- function(g){
  if(!dir.exists('figs')){
    dir.create('figs')
  }
  file_name <- paste0(Sys.time())
  ggsave(filename = paste0('figs/', file_name, '.pdf'), g)
  ggsave(filename = paste0('figs/', file_name, '.png'), g)

}
save_figs(g)
```

# Trends in weekly incidence

```{r}
plot_data <-  df %>%
           mutate(location = ifelse(district == 'MAGUDE',
                                    'Magude',
                                    'Control districts')) %>%
           group_by(date, age_group, location) %>%
           summarise(cases = sum(cases),
                     population = sum(population)) %>%
           ungroup %>%
           mutate(pk = cases / population * 1000) %>%
  mutate(location = factor(location,
                           levels = c('Magude',
                                      'Control districts'))) %>%
  mutate(age_group = paste0('Age ', age_group))

g <- 
  ggplot(data =plot_data,
         aes(x = date,
             y = pk)) +
  geom_line(aes(linetype = location,
                color = location),
            alpha = 0.8) +
  facet_wrap(~age_group) +
  labs(x = 'Week',
       y = 'Cases per 1,000',
       title = 'Weekly incidence') +
  scale_linetype_manual(name = '',
                        values = c(1,2)) +
  theme_publication() +
  scale_color_manual(name = '',
                     values = c(grey(0.1), 'black')) +
  geom_vline(xintercept = as.Date('2015-10-10'),
             col = 'black',
             alpha = 0.8) +
  geom_label(data = data.frame(x = as.Date('2015-10-10'),
                              y = 20,
                              label = 'Intervention\nstart'),
             aes(x = x,
                 y = y,
                 label = label),
             size = 2.5)
g
save_figs(g)

```

# Cases as a percentage of total period average, ages 0-4 years

```{r}

# Define date truncation
date_truncate <- function (date_object, level = c("month", "quarter", "year")) 
  {
    if (is.null(level)) {
      stop("You must provide a level argument of either \"month\", \"quarter\" or \"year\".")
    }
    date_object <- as.Date(date_object)
    if (sum(!is.na(date_object)) == 0) {
      return(date_object)
    }
    if (level == "month") {
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y-%m"), "-01"))
      return(return_object)
    }
    if (level == "quarter") {
      q_month <- (((((as.numeric(format(date_object, "%m"))) - 
                       1)%/%3) + 1) * 3) - 2
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y"), ifelse(nchar(q_month[!is.na(return_object)]) == 
                                                                                    2, "-", "-0"), q_month, "-01"))
      return(return_object)
    }
    if (level == "year") {
      return_object <- date_object
      return_object[!is.na(return_object)] <- as.Date(paste0(format(return_object[!is.na(return_object)], 
                                                                    "%Y"), "-01-01"))
      return(return_object)
    }
  }

# Get a date helper for standardizing dates
wh <- create_date_helper()
wh$month <- as.numeric(format(wh$date, '%m'))

# MANHICA OPD ########################################
load('misc_data/opd_cleaned.RData')
opd_manhica <- opd %>%
  filter(malaria) %>%
  filter(date >= min(bes$date)) %>%
  mutate(age_group = ifelse(age < (5 * 12), '0-4',
                            ifelse(age >= (5 * 12), '5+',
                                   NA))) %>%
  dplyr::select(age_group, date) %>%
  mutate(year = as.numeric(format(date, '%Y')),
         month = as.numeric(format(date, '%m')),
         week = as.numeric(format(date, '%U')) + 1) %>%
  group_by(year, week, age_group) %>%
  summarise(cases = n()) %>%
  ungroup %>%
  mutate(district = 'MANHICA',
         source = 'OPD')
opd_manhica_under5 <-
  opd_manhica %>%
  filter(age_group == '0-4') %>%
  group_by(year, week, district, source) %>%
  summarise(cases = sum(cases, na.rm = TRUE))

# MAGUDE OPD ###########################################
# Read in old OPD Magude data, sent by Bea
old_opd <- readstata13::read.dta13('misc_data/opd_magude_old/survmag_rrs_dis.dta')
old_opd <-
  old_opd %>%
  dplyr::select(yr,
                week,
                mal_hf) %>%
  rename(year = yr,
         cases = mal_hf)

opd_magude_all <- old_opd
opd_magude_all$district <- 'MAGUDE'
opd_magude_all$source <- 'OPD'
rm(old_opd)

# Combine all sources
comparison_magude <-
  opd_magude_all %>%
  bind_rows(bes %>%
              filter(district == 'MAGUDE',
                     disease == 'MALARIA') %>%
              group_by(year, week, district) %>%
              summarise(cases = sum(cases, na.rm = TRUE)) %>%
              ungroup %>%
              mutate(source = 'BES')) %>%
  left_join(wh) %>%
  group_by(year, month, source, district) %>%
  summarise(date = first(date),
            cases = sum(cases, na.rm = TRUE))

comparison_manhica <-
  opd_manhica_under5 %>%
  bind_rows(bes %>%
              filter(district == 'MANHICA',
                     age_group == '0-4',
                     disease == 'MALARIA') %>%
              group_by(year, week, district) %>%
              summarise(cases = sum(cases, na.rm = TRUE)) %>%
              ungroup %>%
              mutate(source = 'BES')) %>%
  left_join(wh) %>%
  group_by(year, month, source, district) %>%
  summarise(date = first(date),
            cases = sum(cases, na.rm = TRUE))

g1 <- ggplot(data = comparison_magude,
       aes(x = date,
           y = cases,
           color = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Magude: all cases',
       x = 'Date',
       y = 'Cases') +
  theme_cism() +
  scale_color_manual(name = 'Source',
                     values = c('darkgreen', 'darkorange'))

g2 <- ggplot(data = comparison_manhica,
       aes(x = date,
           y = cases,
           color = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Manhica: 0-4 only',
       x = 'Date',
       y = 'Cases') +
  theme_cism() +
  scale_color_manual(name = 'Source',
                     values = c('darkgreen', 'darkorange'))

# Comparison manhica
x <- 
  comparison_manhica %>%
  ungroup %>%
  group_by(district, source) %>%
  mutate(avg = mean(cases)) %>%
  ungroup %>%
  mutate(y = cases / avg * 100)
g <- 
  ggplot(data = x %>%
  filter(date <= '2016-08-30'),
       aes(x = date,
           y = y, 
           linetype = source)) +
  geom_line(alpha = 0.8) +
  labs(title = 'Cases as a percentage of total period average',
       subtitle = 'Manhiça district, ages 0-4 only',
       x = 'Date',
       y = 'Percent') +
  theme_publication() +
  scale_linetype_manual(name = '',
                     values = c(1, 2)) 
g
save_figs(g)
```

```{r}
# Zip up all figures
zip('figs/figures.zip',
    files = paste0('figs/', dir('figs')))
```

```{r eval=FALSE}
file.edit(
  tufte:::template_resources(
    'tufte_html', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```

### 2017-09-06 reports/aes/aes_effectiveness.Rmd
update

---
title: "Short-term impact of a malaria elimination campaign in Southern Mozambique"
subtitle: Evidence from a quasi-experimental approach applied to routine surveillance data
author: "AES, Barcelona, 2017"
header-includes:
  - \widowpenalties 1 150
fig_height: 5
fig_width: 7
bibliography: bibliography.bib
output:
  beamer_presentation:
    theme: "Berlin"
    colortheme: "seahorse"
    fonttheme: "structuresmallcapsserif"
    toc: true
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
library(knitr)
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r, results = 'hide'}
# Packages
library(ggplot2)
library(cism)
library(rgeos)
library(maptools)
library(rgdal)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(leaflet)
library(readr)
library(ggthemes)
library(ggrepel)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html") 
# Define a theme
theme_maltem <- ggthemes::theme_hc
```

```{r}
source('../../master.R', chdir = TRUE)

tablify <- function(x, n = 5){
  DT::datatable(x,
                selection = 'none',
                escape = FALSE,
                options = list(#sDom  = '<"top">lrt<"bottom">ip',
                               pageLength = n, 
                               dom = 'tip'))
}
kablify <- function(x, size = 16){
  kable(x) %>% 
  kable_styling(bootstrap_options = c("striped",
  "hover",
  "condensed"),
  font_size = size,
  position = 'float_left')
}

# ## 
# 
# <img src="img/tree.jpg" alt="img/tree.jpg" style="width: 600px; align="center"/>

# # Synthetic control reproduction
# 
# # Get packages
# library(Synth)
# 
# # Prepare data
# gen_dummy <- function(condition){ifelse(condition, 1, ifelse(!condition, 0, NA))}
# 
# model_data <- 
#   df %>%
#   filter(age_group == '0-4') %>%
#   filter(disease == 'MALARIA',
#          district != 'XAI-XAI CITY',
#          district != 'MANHICA',
#          district != 'MASSINGIR') 
# 
# # Get a standardized time variable
# model_data <-
#   model_data %>%
#   mutate(time = as.Date(paste0(year, '-01-01')) + (week*7)) %>%
#   mutate(time = as.numeric(time)) %>%
#   mutate(high_season = ifelse(month %in% 1:3, 1, 0))
# 
# # Generate a oct-dec lag
# high_season <- model_data %>%
#   filter(high_season == 1) %>%
#   group_by(district, year) %>%
#   summarise(cases = mean(pk)) %>%
#   ungroup %>%
#   group_by(district) %>%
#   mutate(cases_lag = dplyr::lag(cases, n = 1, default = NA)) %>%
#   ungroup %>%
#   dplyr::select(-cases)
# model_data <-
#   left_join(model_data,
#             high_season)
# # Get 1 week lags too
# model_data <-
#   model_data %>%
#   arrange(date) %>%
#   group_by(district) %>%
#   mutate(cases_lag1 = dplyr::lag(pk, n = 1, default = NA),
#          cases_lag2 = dplyr::lag(pk, n = 2, default = NA))
# 
# # Filter away some stuff
# model_data <- 
#   model_data %>%
#   filter(week < 52) %>%
#   filter(date >= '2013-10-01')
# 
# # Get data with dummies
# model_formula <- lm(pk ~  high_season + year + time, data = model_data)
# model_matrix <- data.frame(model.matrix(model_formula)) %>%
#   dplyr::select(-`X.Intercept.`) %>%
#   mutate(pk = model_data$pk,
#          district = model_data$district,
#          itn_coverage = model_data$itn_coverage) %>%
#   mutate(district_name = district) %>%
#   mutate(district = as.numeric(factor(district)),
#          cases_lag = model_data$cases_lag,
#          cases_lag1 = model_data$cases_lag1,
#          cases_lag2 = model_data$cases_lag2)
# 
# left <- 
#   expand.grid(district = sort(unique(model_matrix$district)),
#               time = sort(unique(model_matrix$time)))
# right <- 
#   model_matrix %>%
#   group_by(district, time) %>%
#   tally
# balanced <- left_join(left, right)
# model_matrix <-
#   left_join(left, model_matrix)
# 
# # Fill the NAs
# model_matrix <-
#   model_matrix %>%
#   arrange(time) %>%
#   group_by(district) %>%
#   mutate(high_season = zoo::na.locf(high_season), 
#          year = zoo::na.locf(year),
#          pk = zoo::na.locf(pk),
#          itn_coverage = zoo::na.locf(itn_coverage),
#          district_name = zoo::na.locf(district_name))
# 
# model_matrix <- data.frame(model_matrix)
#   
# prior_time <- sort(unique(model_matrix$time))
# prior_time <- prior_time[prior_time <= as.numeric(as.Date('2015-08-01'))]
# prepared_data <- 
#   Synth::dataprep(foo = model_matrix,
#                   predictors = c('itn_coverage',
#                                  'cases_lag',
#                                  'cases_lag1',
#                                  'cases_lag2'),
#                   predictors.op = 'mean',
#                   dependent = 'pk',
#                   time.variable = 'time',
#                   unit.variable = 'district',
#                   treatment.identifier = which(sort(unique(model_data$district)) == 'MAGUDE'),
#                   controls.identifier = which(sort(unique(model_data$district)) != 'MAGUDE'),
#                   unit.names.variable = 'district_name',
#                   time.predictors.prior = prior_time,
#                   time.optimize.ssr = prior_time,
#                   time.plot = sort(unique(model_matrix$time)))
# 
# out <- Synth::synth(data.prep.obj = prepared_data)
# # synth_tab <- synth.tab(synth.res = out,
# #           dataprep.res = prepared_data,
# #           round.digit = 3)
# # # Predictor weights
# # out$solution.v
# # # Unit weights
# # out$solution.w
```

# Introduction

## The burden of malaria

- 2015: 212M cases; 429K deaths (global). 
- Mozambique among ten highest burden countries in world. 
- 29% reduction since 2010.  
- Now, focus on _elimination_ and _eradication_.

## How to eliminate

- Mass drug administration (MDA).  
- Other, more time-tested strategies (ITNs, IRS).
- The need for operational research.

## It's hard to evaluate elimination strategies

- Difficult to measure impact of interventions in operational research.  
- OR is usually not experimental, not randomized, not controlled.
- Before vs. after and case vs. control approaches are useful, but biased.

## Why evaluate?

- The WHO explicitly supports “the need for more research on the optimum methods of implementing MDA programs, promoting community participation and compliance with treatment, and evaluating their effectiveness.

- As elimination initiatives imply both strong financial/economic efforts and a high risk of failure, their affordability, sustainability and cost-effectiveness are among the main concerns of funders, stakeholders and policymakers

- Evidence of the causal impact of malaria elimination initiatives is a pre-requisite for evaluating their cost-effectiveness. Finally, rigorous impact evaluations of malaria elimination programs are essential for sustained financial and political support.

## How to evaluate?

- Since we don't have an experiment, we need quasi-experimental approaches.  
- The synthetic control method (SCM).  
- We want to construct a _realistic_ counterfactual telling us _what would have happened_ without the campaign.

## What are we evaluating, exactly?

- Malaria elimination campaign in Magude.  
- Deployed during rainy seasons of 2015-6 and 2016-7. 
- Consisted of MDA, IRS, and surveillance. 
- ITN coverage was already high from 2014.

# Methods

## Data

- Mozambique’s Boletim Epidemiológico Semanal (BES) of the NMCP.  
- Weekly clinical cases, by age group, by health facility, for each of a few high risk diseases
, the epidemic disease reporting system used by the NMCP. 
- Population estimates from the National Statistical Institute (INE).

## Identification strategy

- Treatment group is district of Magude, Maputo province. 
- Control group consists of the other districts in Maputo and Gaza province. 
- The selection of control districts from locations around Magude is to ensure treatment and control groups have similar epidemiological characteristics.

## Co-variates

Our analysis uses variables as predictors of malaria incidence:

- LLINs

- Average weekly temperature

- Weekly rainfall

## Rainfall

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

g <- ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = precipitation,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = precipitation,
           group = district)) +
  labs(x = 'Date',
       y = 'Milimeters',
       title = 'Precipitation: Magude vs. other districts',
       subtitle = 'Monthly average precipitation, Magude in red') +
  theme_maltem()
g 
```

## Temperature

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(place = factor(place, levels = c('Other districts', 'Magude'))) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, district, place) %>%
  summarise(precipitation = mean(precipitation),
            temp = mean(temp),
            temp_max = max(temp_max),
            temp_min = min(temp_min))

ggplot() +
  geom_line(alpha = 0.2,
            data = x %>% filter(place == 'Other districts'),
       aes(x = date,
           y = temp,
           group = district)) +
    geom_line(alpha = 0.8,
              color = 'darkred',
            data = x %>% filter(place == 'Magude'),
       aes(x = date,
           y = temp,
           group = district)) +
  labs(x = 'Date',
       y = 'Degrees (celcius)',
       title = 'Average temperature: Magude vs. other districts',
       subtitle = 'Monthly average, Magude in red') +
  theme_maltem() +
  xlim(as.Date('2014-01-01'),
               max(x$date))
```

## Difference in differences

- We _could_ compare changes in malaria incidence to other districts using a difference-in-differences (DD) model. 
- Often used to estimate treatment effects in such settings by contrasting changes in outcome pre- and -post intervention, for the treatment and control groups.
- Eliminates confounding due to time-invariant differences between treatment and control groups, arising from unobservable factors that might influence malaria incidence. 
- Assumes that any macro shocks such as time effects are common to the groups under evaluation.

## Why we can't use difference in differences

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

g1 <- ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Monthly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))

x2 <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  mutate(date = as.Date(paste0(year, '-', month, '-01'))) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  group_by(date) %>%
  summarise(magude_to_other = p[place == 'Magude'] / 
              p[place == 'Other districts'] * 100)

g2 <- ggplot(data = x2,
       aes(x = date,
           y = magude_to_other)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  geom_hline(yintercept = 100,
             col = 'darkred',
             alpha = 0.6,
             lty = 2) +
  labs(x = 'Date',
       y = 'Percent') +
  labs(title = 'Magude incidence as % of other districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  xlim(as.Date('2013-01-01'), max(x$date))
Rmisc::multiplot(g1, g2, cols = 2)
```

## Why not DiD?

- Magude shows a downward trend in incidence over time, even prior to the elimination campaign.
- Lack of a visible parallel trend between treatment and control groups

## SCM

- We employ the synthetic control method (SCM).  
- Weighted combination of potential control districts, the synthetic control, using the districts that best approximate the most relevant characteristics of malaria incidence in Magude before the intervention. 
- Characteristics include pre-intervention malaria incidence during rainy season, temperature and and rainfall.
- Good match based on RMSE.

## Advantages of SCM

- Makes explicit the relative contribution of selected control districts to the counterfactual; the similarities between Magude and the synthetic control can be evaluated by comparing pre-intervention outcomes and predictors.  
- Allows the effects of unobserved factors on malaria incidence to vary with time, relaxing the parallel trend assumption of DD discussed earlier.

## Limitations of SCM

- Does not allow assessing the significance of the results using commonly used (large-sample) inferential techniques. 
- Instead, we use placebo experiments based on permutation techniques to make inference:  sequentially applying the SCM assuming each control district was “treated” to generate a distribution of placebo effects, and compare Magude's effect with these distributions.

# Results

## All districts' incidences (aggregated age groups)

```{r}
x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p)) +
  geom_line(alpha = 0.8,
            color = 'darkred') +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'All age groups') +
  theme_maltem() +
  facet_wrap(~district) +
  theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))
```

## All districts' incidences (separated by age groups)

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  group_by(date, age_group, district) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = age_group,
           color = age_group)) +
  geom_line(alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('darkorange', 'darkgreen')) +
  labs(title = 'Weekly incidence: all districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~district) +
  # theme(legend.position="none") +
  theme(strip.background = element_rect(fill="white"),
        strip.text = element_text(color = grey(0.1), size = 6)) +
  theme(axis.text.x = element_text(angle = 90))
```

## Magude vs. other districts

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'All age groups') +
  theme_maltem()
```

## Magude vs. other districts (separated by age group)

```{r}

x <- df %>%
  filter(disease == 'MALARIA') %>%
  mutate(place = ifelse(district == 'MAGUDE', 'Magude',  'Other districts')) %>%
  group_by(date, place, age_group) %>%
  summarise(p = sum(cases) / sum(population) * 1000)

ggplot(data = x,
       aes(x = date,
           y = p,
           group = place)) +
  geom_line(aes(color = place), alpha = 0.8) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2010:2017, '-01-01'))),
             color = 'black',
             alpha = 0.3) +
  labs(x = 'Date',
       y = 'Weekly cases per 1,000') +
  scale_color_manual(name = 'District',
                     values = c('red', 'black')) +
  labs(title = 'Weekly incidence: Magude vs. other districts',
       subtitle = 'Segregated by age group') +
  theme_maltem() +
  facet_wrap(~age_group)
```

## Our weighted synthetic control

```{r}
uw <-
  df %>%
  filter(age_group == '0-4') %>%
  filter(disease == 'MALARIA',
         district != 'XAI-XAI CITY',
         district != 'MANHICA',
         district != 'MASSINGIR') %>%
  group_by(district) %>%
  tally
uw$w.weight <-
  c(0.00,
    0.22,
    0.00,
    0.19,
    0.01,
    0.00,
    0.00,
    0.27,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.07,
    0.00,
    0.00,
    0.24
    )
uw$age_group <- '0 to 4'

older <- uw
older$age_group <- '5+'
older$w.weight <- 
  c(0.00,
    0.00,
    0.00,
    0.00,
    0.10,
    0.00,
    0.00,
    0.26,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.00,
    0.35,
    0.29
)
uw <- bind_rows(uw, older)
# Overwrite with Ranjeeta's weights!
ggplot(data = uw,
       aes(x = district,
           y = w.weight)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkorange') +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = 'District',
       y = 'Weight',
       title = 'Weights by district') +
  ggthemes::theme_hc() +
  facet_wrap(~age_group,
             nrow = 2)
```

## Is it a good match?

```{r}
# Copy of Ranjeeta's table
x <- dplyr::data_frame(Indicator = c('Incidence, rainy season 2015',
                              'Incidence, rainy season 2014', 
                              'LLIN coverage',
                              'Pre-intervention precipitation',
                              'Pre-intervention temperature',
                              'Post-intervention precipitation'),
                `Magude 0-4` = c(3.65, 8.87, 24.82, 1.71, 24.52, 1.84),
                `Synthetic Magude 0-4` = c(5.1, 6.93, 24.77, 1.74, 24.51, 1.82),
                `Magude 5+` = c(4.02, 8.53, 24.83, 1.7, 24.52, 1.84),
                `Synthetic Magude 5+` = c(4.12, 6.29, 25.22, 1.71, 24.49, 1.83))
```

\begin{tiny}

```{r}

knitr::kable(x, format = 'latex')
```

\end{tiny}

## 0-4 year-olds: Magude vs synthetic control

<!-- <img src="img/synthetic1.png" alt="img/synthetic1.png" style="width: 800px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/synthetic1}

## 5+ year-olds: Magude vs synthetic control

<!-- <img src="img/synthetic2.png" alt="img/synthetic2.png" style="width: 800px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/synthetic2}

## Effect of interevention

Relative to the synthetic control, the average treatment effect during the first post-intervention malaria season was a reduction in weekly incidence by 4 and 2 cases per 1,000 in those aged 0-4 and 5 or older, respectively. During peak malaria season (January - March), 1,939 cases were prevented in 2016, and 3,405 in 2017.

<!-- <img src="img/ranjeeta_table.png" alt="img/ranjeeta_table.png" style="width: 600px; align="center"/> -->

\includegraphics[width=0.9\textwidth]{img/ranjeeta_table}

This represents a 91.29% and 81.3% reduction in the incidence of malaria relative to the hypothetical control.

# Discussion

## Our findings

-  Across age groups during the peak malaria season (January-March) incidence declined by 91.29% and 81.3% in 2016 and 2017, respectively.  
- Has not achieved zero cases.  
- Unsure, at this point, to what extent we can attribute cases to importation or other causes.

## Comparison with other findings

- Similar findings obtained in a cluster randomized controlled trial (RCT) estimating the effectiveness of MDA with DHA/PPQ in Zambia.

- In that study, while the number of malaria cases significantly declined after MDA in low-transmission areas, drop was not significant in areas of high transmission (ie, Magude).

- Also consistent with the predictions of a consensus study that put together 4 models focusing on the impact of MDA with a drug having similar characteristics as DHA/PPQ.

- The latter predicted an immediate (but transient w/o other interventions) decline in malaria prevalence after MDA.

## Quasi-experimental methods in similar contexts

- Quasi-experimental approaches have been applied in other elimination campaigns.

- Barofsky et al. used SCM as an approach for the identification of the impact of a malaria campaign in a district in Uganda.   
- These studies are from 50s-60s and look at economic indicators, such as primary school completion or households consumption.   
- To our knowledge, this is the first quasi-experimental evaluation of an ongoing elimination initiative using routine national surveillance data on health indicators.

## Comparison with RCT

- Randomized control trials (RCTs) are the gold standard for impact evaluations.   
- However, the implementation of RCTs may imply financial, practical and ethical challenges.  
- Furthermore, RCTs have debatable usefulness in identifying intervention feasibility and implementation bottlenecks, which are particularly important in disease elimination contexts.
- Quasi-experimental approaches are particularly useful when applied to routinely collected (administrative/observational) data.

## Scaling up the method?

- Given the focus over the last decade on malaria elimination with both large and small scale campaigns across the globe – quasi-experimental approaches offer a rigorous and inexpensive alternative to randomized-control trials.

- Observational data is widely (and often publicly) available and true experimental approaches are often not feasible:  quasi-experimental approaches like the SCM are an excellent alternative for the evaluation of malaria elimination initiatives.

- Approach is not only suitable for the estimation of the impact of the intervention in question, but also generalizable to malaria elimination campaigns globally.

## Limitations

1. We rely on surveillance malaria cases reported at the health facility level, which is an incomplete and imperfect proxy for true malaria infection incidence since health-seeking behavior is especially low in low-income countries settings and may be different across geography and time.

2. The health facility registries from which we gathered data are themselves imperfect; we have no mechanisms by which to validate the entirety of the data. Still, BES data offers the best country-wide comparison of malaria trends at a granular level (weekly and by district).

3. Even with quasi-experimental data, we still can't fully control for abnormalities... (see next slide)

## 2017 has been anything but normal

```{r}
x <- 
  df %>%
  group_by(year, month) %>%
  summarise(p = sum(cases) / sum(population) * 1000) %>%
  ungroup %>%
  mutate(year = factor(year))

cols <- colorRampPalette(RColorBrewer::brewer.pal(n = 9, 'Spectral'))(length(unique(x$year)))
ggplot(data = x,
       aes(x = month,
           y = p,
           group = year,
           color = year)) +
  geom_line() +
  geom_point() +
  scale_color_manual(name = 'Year',
                     values = cols) +
  labs(x = 'Month',
       y = 'Average incidence',
       title = 'Incidence by month, all districts',
       subtitle = 'Wide variability in monthly incidence') +
  theme_maltem()

```

## Implications

- The strategy reduced malaria significantly.   
- Effectiveness quantified accurately enough to allow for granular cost-effectiveness.   
- Synthetic control method useful and generalizable for the estimation of similar initiatives’ effectiveness.   
- Policy-makers, researchers, and those actively engaged in or planning disease elimination initiatives, can benefit from both the methods and findings from this study.

## Acknowledgments

- NMCP for releasing BES data.  
- CISM for support.  
- Bill and Melinda Gates Foundation and “La Caixa” Foundation for funding MALTEM.   
- JB is funded by the Trans Global Health doctorate program.

### 2017-09-06 scm.Rmd
update

--- 2017-08-18 paper.Rmd
@@
 ---
-title: "Cost-effectiveness of a malaria elimination campaign"
+title: "Effectiveness of a malaria elimination campaign"
 subtitle: "Cross-software reproducibility check"
-author: "Joe Brew, Laia Cirera, Ranjeeta Thomas, Elisa Sicuri"
 output:
